---
title: "Shelton et al. Quant eDNA Model"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

**Date written: May 20 2023**

**Author: M Fisher**

**Last Run: `r Sys.Date()`**


# Description

Run the Shelton et al. (2023) quantitative eDNA model to generate alpha values for all species in the mock communities. 

A note about error messages, from Ryan Kelly:

- Max Tree Depth is a warning you can ignore.  It just means sampling wasn’t as efficient as it could have been, but the results are still totally valid.
- Divergent transitions are intermediate-serious warnings.  These mean that sampling isn’t working quite right, but if there are small numbers of divergent transitions, that’s not too bad.
- Rhat warnings are serious and mean that the model isn’t converging; something is broken.
- Low sample-size warnings aren’t serious; you can either ignore them or else run the chains for longer.


*Run notes:*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(here)
library(ggplot2)
library(ggrepel)
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

source(here("../../","R","calibrate_metabarcoding_RPK.R"))

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

n_pcr_samples <- read_csv(here('../../','data','qm_model','qm_model_SAMPLEpcr.csv'))

set.seed(555)
```


## Alpha Run

This is the data produced for estimating amplification efficiencies of all species in the mock communities, not just the ones in the diet data. 

```{r}
load(here('qm_model_dataLIST_matchAlpha_2023-8-29.rds'))
EGC_obs <- EGC_obs2

rm(EGC_obs2)
```


I forgot to make sure that we get an amplification efficiency for *Cottus asper*, so let's just pick another fish in the diet data set and replace it with *Cottus*.

```{r}
EGC_obs[[2]] %<>% mutate(species=ifelse(species=="Pholis ornata", "Cottus",species))
```



### re-format data

insert missing zeros, and rename the levels of the data to match eDNA naming conventions.
```{r}
EGC_obs[[1]] <- EGC_obs[[1]] %>% 
  unite(Community, CommType, tech, col = "Sample", remove = T, sep = "#") %>% 
  pivot_wider(names_from = Sample, values_from = Nreads, values_fill = 0) %>% 
  pivot_longer(-c(species, totReads, propReads, b_proportion, N_pcr_mock), names_to = "Sample", values_to = "Nreads") %>% 
  separate(Sample, into = c("Community", "CommType", "tech"), sep = "#")

EGC_obs[[2]] <- EGC_obs[[2]] %>% 
  unite(station, creek, time, tech, col = "Sample", remove = T, sep = "#") %>% 
  pivot_wider(names_from = Sample, values_from = Nreads, values_fill = 0) %>% 
  pivot_longer(-c(species), names_to = "Sample", values_to = "Nreads") %>% 
  separate(Sample, into = c("station", "creek", "time", "tech"), sep = "#")
```

Filter samples according to a minimum read-depth cutoff of n=100
```{r}
#plot the read depths going into the quant model?
png(here('qm_model_sample-read-depth_matchAlphas_2023-8-29.png'))
EGC_obs[[2]] %>%
  filter(species %in% EGC_obs[[1]]$species) %>%                # species in quant model
  group_by(station,tech) %>% summarise(total_reads=sum(Nreads)) %>%
  ggplot(aes(x=total_reads)) + geom_histogram() + 
  labs(x="Read Depth into Model",y="Number of Crab+Replicates") +theme_bw()
dev.off()

#filter for > 100 reads?
low_read_crab <- EGC_obs[[2]] %>%
  filter(species %in% EGC_obs[[1]]$species) %>% 
  group_by(station,tech) %>% summarise(total_reads=sum(Nreads)) %>% filter(total_reads < 100)
low_read_crab
```
```{r}
EGC_obs[[2]] <- EGC_obs[[2]] %>% anti_join(dplyr::select(low_read_crab,station,tech))
```


### Run for all samples

Since we're not using the posterior Bayes estimates of sample DNA proportions, we don't have to split the data frames according to the number of PCR cycles for each sample. We can also just make up a single number for the number of pcr cycles argument in the `makeDesign` function


create input matrix
```{r}
my_dDNA_data <- format_metabarcoding_data(EGC_obs[[2]], N_pcr_mock=40,
                                          EGC_obs[[1]],
                                          Level_1_treatment_envir <- "station",
                                          Level_2_treatment_envir <- "tech",
                                          Level_3_treatment_envir <- NA,
                                          Level_1_treatment_mock <- "Community",
                                          Level_2_treatment_mock <- "tech",
                                          Level_3_treatment_mock <- NA)



mydata <- makeDesign(my_dDNA_data, N_pcr_cycles = 40)
```
check that all species are present
```{r}
mydata$sp_list
```

run model
```{r}
QM_Bayes_alpha <- QM_bayes("quant_metabar_multinom.stan", mydata, NCHAINS = 3, ITER = 5000 )
saveRDS(QM_Bayes_alpha,file="QM_Bayes_out_100rmin_matchAlphas_2023-08-29.rds")
```
Warning: There were 13500 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 12. See
https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceededWarning: Examine the pairs() plot to diagnose sampling problems



What are the alphas?
```{r}
QM_Bayes_alpha$Bayes_alpha_est
```



### Save 

first, the input data frame
```{r}
write_csv(QM_Bayes_alpha$Bayes_alpha_est,here('../../','data','qm_model','QM_Bayes_out_AlphaRun_alpha_est.csv'))
```



