---
title: 'Fig 4: Mixture Model Results'
author: "M Fisher"
date: "2023-09-01"
output: html_document
---

## Description

Show quant mixture model / zoid results
  (a) tracks percent reads through all steps after blast
  (b) shows what proportion of all prey reads are represented by 
            the calibrated prey species
  (c) shows the zoid "average crab" proportions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magrittr)
library(here)
library(grid)
library(cowplot)

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

filter_colors_df <- data.frame(filter=c("total reads","missing","predator","bait","non-target, non-specific","prey","calibrated","leftover"),
                               color=c("black",gray.colors(n=1,start=0.2,end=0.2),"darkgreen",gray.colors(n=2,start=0.55,end=0.7),"lightblue","darkorange","gray80"))

filter_colors_df2 <- data.frame(filter=c("total reads","missing","predator","bait","non-target, non-specific","prey","calibrated"),
                               color=c("black",gray.colors(n=1,start=0.2,end=0.2),"darkgreen",gray.colors(n=2,start=0.55,end=0.7),"lightblue","darkorange"))


```
```{r}
miseqruns=c(2,3,4,5)
```

## Part 1: track reads

### Data 

quant model input
```{r}
load(here('../','data','zoid',"qm_model_dataLIST_2023-8-29.rds"))
```


quant model output
```{r}
QM_Bayes_filter <- read_csv(here('../','data','zoid','QM_Bayes_posterior_means_filtered_08-29-2023.csv'))

keep_crab <- QM_Bayes_filter %>% filter(name!="Carcinus maenas") %>% pull(crab)
```

dDNA data:


*all prey in final data set*
```{r}
tax.prey.clean <- read_csv(here::here('../','data','results','allRuns_BF3_filtered_FINAL_unique_taxa.csv'))
tax.prey <- read_csv(here::here('../','data','results','allRuns_BF3_filtered_unique_taxa.csv'))
```

*record of ASV filtering for all that matched to BLAST*
Raw BLAST output: matches to database**
```{r blast}
miseqruns <- unique(tax.prey$MiSeqRun)
for(r in miseqruns){
  if(r==miseqruns[1]){
    tax.unfiltered <- read_csv(here('../','data','blast',paste0('run',r,'_bf3_sample_CEGblast_lca_taxonomy.csv'))) %>% mutate(sample_id=as.character(sample_id))
    tax.unfiltered %<>% mutate(MiSeqRun=r)
  } else{
    tax.unfiltered <- tax.unfiltered %>% 
      bind_rows(read_csv(here(here('../','data','blast',paste0('run',r,'_bf3_sample_CEGblast_lca_taxonomy.csv')))) %>% mutate(sample_id=as.character(sample_id)) %>% mutate(MiSeqRun=r))
  }
}

tax.unfiltered %<>% dplyr::select(-dataset)

# grab controls
tax.controls <- tax.unfiltered %>% filter(sample_id %in% c("WACO21-ENegative","WACO21-PCRNegative","WACO21-Positive","93","PositiveControl","NegativeControl"))
  
# just samples we're interested in
tax.unfiltered.sub <- tax.unfiltered %>% left_join(dplyr::select(tax.prey,Sample_name,sample,tech,MiSeqRun) %>% distinct(), by=c("Sample_name","MiSeqRun")) %>%
  filter(sample %in% keep_crab)
```

Check what IDs are present for *Mya* genus
```{r}
filter(tax.unfiltered.sub, genus=="Mya") %>% dplyr::select(sample,tech,taxon,rank) %>% distinct()
```

And in our quant model output, how is *Mya* represented?
```{r}
"Mya arenaria" %in% QM_Bayes_filter$name 
"Mya" %in% QM_Bayes_filter$name 
```



### Track reads

Get total reads
```{r}
total.reads.df <- tax.unfiltered.sub %>%
  group_by(sample,tech) %>% summarise(total_reads=sum(nReads))
```


Get missing reads 
```{r}
tax.missing.summary <- tax.unfiltered.sub %>%
  filter(is.na(taxon)) %>%
  group_by(sample,tech) %>% summarise(Treads=sum(nReads)) %>%
  mutate(step="missing")
```

Get control reads
```{r}
tax.controls.summary <- tax.unfiltered.sub %>%
  filter(Hash %in% tax.controls$Hash) %>%
  group_by(sample,tech) %>% summarise(Treads=sum(nReads)) %>%
  mutate(step="control")
```


Get predator reads
```{r}
tax.egc <- tax.unfiltered.sub %>% filter(taxon %in% c("Carcinus maenas","Carcinus"))

tax.egc.summary <- tax.egc %>% group_by(sample,tech) %>%
  summarise(Treads=sum(nReads)) %>%
  mutate(step="predator")
```

Breaking down other reads
```{r}
other.tax <- tax.unfiltered.sub %>%
  filter(!(is.na(taxon))) %>%  # missing
  filter(!(taxon %in% c("Carcinus maenas","Carcinus"))) %>% # predator
  filter(!(Hash %in% tax.controls$Hash)) %>%    # control
  mutate(step=ifelse(Hash %in% tax.prey$Hash & taxon %in% tax.prey.clean$taxon, "prey",
                     ifelse(grepl("Scomber",taxon),"bait", "non-target, non-specific"))) %>%
  mutate(step=ifelse(taxon %in% EGC_obs[[1]]$species | taxon=="Mya arenaria", "calibrated",step))

other.tax.summary <- other.tax %>% group_by(sample,tech,step) %>%
  summarise(Treads=sum(nReads))
```

just check that Mya / Mya arenaria is marked as calibrated
```{r}
other.tax %>% filter(taxon %in% c("Mya","Mya arenaria")) %>% pull(step)
```

### Read distribution / change by step

Read distribution
```{r echo=FALSE}
all.summary <- bind_rows(tax.missing.summary,tax.controls.summary,tax.egc.summary,other.tax.summary) %>%
  left_join(total.reads.df, by=c("sample","tech")) %>%
  mutate(pReads=Treads/total_reads)
```


```{r echo=FALSE}
all.summary$step <- factor(all.summary$step,levels=c("missing","control","bait","non-target, non-specific","predator","prey","calibrated"))
ggplot(all.summary,aes(x=step,y=pReads, col=step)) + 
  geom_boxplot()  + 
  geom_jitter() + 
  scale_color_manual(values=filter_colors_df$color) +
  labs(y="Proportion of Sample Reads") + theme(legend.position="none")
```


Read change by step

```{r echo=FALSE}
all.sequential <- total.reads.df %>%
  left_join(tax.missing.summary, by=c("sample","tech")) %>%
  mutate(missing=total_reads-Treads) %>% 
  dplyr::select(sample,tech,total_reads,missing) %>%
  left_join(tax.egc.summary) %>%  mutate(Treads=ifelse(is.na(Treads), 0, Treads)) %>%
  mutate(predator=missing-Treads) %>%
  dplyr::select(sample,tech,total_reads,missing,predator) %>%
  left_join(other.tax.summary %>% pivot_wider(names_from=step,values_from=Treads, values_fill=0), by=c("sample","tech")) %>%
  mutate(bait=predator-bait) %>%
  mutate(`non-target, non-specific`=bait-`non-target, non-specific`) %>%
  mutate(prey=`non-target, non-specific`-prey) %>%
  mutate(leftover=prey-calibrated) %>%
  mutate_at(c("missing","predator","bait","non-target, non-specific", "prey", "calibrated","leftover"), ~.x/total_reads) %>%
  mutate(total_reads=1) %>% rename(`total reads`=total_reads) %>%
  pivot_longer(cols=3:10, names_to="step",values_to="prop_reads_remaining")
```

```{r echo=FALSE}
all.sequential$step <- factor(all.sequential$step,levels=c("total reads","missing","predator","bait","non-target, non-specific","prey","calibrated","leftover"))
ggplot(all.sequential,aes(x=step,y=prop_reads_remaining, col=step)) + 
  geom_boxplot()  + 
  geom_jitter() + 
  scale_color_manual(values=filter_colors_df$color) +
  labs(y="Proportion of Sample Reads Remaining",x="Filtering Step") + theme(legend.position="none")
```


Save the data frame, for making figure 1
```{r}
if(!knitting){
  
write_csv(all.sequential %>% filter(step!="leftover"), here('../','data','zoid','zoid_track_reads_calibrated_crab.csv'))
  
}
```


### Prey Reads 

Zoom in on what proportion of all prey identified (% unique taxa, % reads) are represented by the calibrated prey.
```{r echo=FALSE}
checked.prey <- tax.prey.clean %>%
  mutate(calibrated=ifelse(taxon %in% EGC_obs[[1]]$species | taxon == "Mya arenaria", "yes", "no"))

## proportion taxa / reads
checked.prey.summary <- checked.prey %>%
  mutate(site_type=ifelse(grepl("Oysterville",site_month) | grepl("Long Beach",site_month), "clam bed","slough")) %>%
  group_by(sample) %>%
  mutate(sample.taxa=length(unique(taxon)),sample.reads=sum(nReads)) %>%
  ungroup() %>%
  group_by(site_type,sample,calibrated,sample.taxa,sample.reads) %>%
  summarise(n.taxa=length(unique(taxon)),n.reads=sum(nReads)) %>%
  mutate(pTaxa=n.taxa/sample.taxa,
         pReads=n.reads/sample.reads) %>%
  filter(grepl("MC",sample)==FALSE)
```


Proportion taxa / reads overall:
```{r echo=FALSE}
checked.prey.summary %>% filter(calibrated=="yes") %>%
  pivot_longer(cols=c("pTaxa","pReads"), names_to="metric",values_to="proportion") %>%
  mutate(metric=ifelse(metric=="pTaxa","Unique Taxa","Reads")) %>%
  ggplot(aes(x=metric,y=proportion)) +
  geom_boxplot(color="darkorange") + 
  geom_jitter(alpha=0.6,color="darkorange") +
  labs(x="",y="proportion prey calibrated per crab") + theme_bw() +
  theme(panel.background=element_rect(color="lightblue"))
```


Proportion taxa / reads by site type:
```{r echo=FALSE}
checked.prey.summary %>% filter(calibrated=="yes") %>%
  pivot_longer(cols=c("pTaxa","pReads"), names_to="metric",values_to="proportion") %>%
  mutate(metric=ifelse(metric=="pTaxa","Unique Taxa","Reads")) %>%
  ggplot(aes(x=metric,y=proportion)) +
  geom_boxplot(color="darkorange") + 
  geom_jitter(alpha=0.6, color="darkorange") +
  facet_wrap(~site_type) +
  labs(x="",y="proportion prey calibrated per crab") + theme_bw() +
  theme(panel.background=element_rect(color="lightblue"),
        strip.background=element_rect(color="lightblue"))
```



Save the data framef or making figure 1
```{r}
if(!knitting){
  
write_csv(checked.prey.summary, here('../','data','zoid','zoid_track_prey_calibrated_crabs.csv'))
  
}
```





## Part 2: Figures

### Data

```{r cars}

track_reads <- read_csv(here('../','data','zoid','zoid_track_reads_calibrated_crab.csv'))

track_prey <- read_csv(here('../','data','zoid','zoid_track_prey_calibrated_crabs.csv'))

fitted_vals_1 <- read_csv(here('../','data','results','zoid_fitted_vals_preyD100_2023-08-29.csv'))

fitted_vals_clambed <- read_csv(here('../','data','results','zoid_fitted_vals_CLAMBED_preyD100_2023-08-29.csv'))

fitted_vals_slough <- read_csv(here('../','data','results','zoid_fitted_vals_SLOUGH_preyD100_2023-08-29.csv'))

fitted_vals_bysite <- fitted_vals_clambed %>% mutate(site_type="Clam Bed") %>%
  bind_rows(fitted_vals_slough %>% mutate(site_type="Slough"))
```



Create key with prey names for figure against scientific names (used in zoid input/output)
```{r}
prey_names <- data.frame(scientific = c("Metacarcinus magister",
                                        "Crangon franciscorum","Cymatogaster aggregata",
                                        "Batillaria attramentaria",
                                        "Mya arenaria","Leptocottus armatus",
                                        "Hemigrapsus oregonensis","Ruditapes philippinarum"),
                         common = c("Dungeness crab\n(C. magister)",
                                    "Sand shrimp\n (C. franciscorum)","Shiner perch\n(C. aggregata)",
                                    "Mud snail\n(B. attramentaria)",
                                    "Soft-shell clam\n(M. arenaria)","Staghorn sculpin\n(L. armatus)",
                                    "Hairy shore crab\n(H. oregonensis)","Manila clam\n(R. philippinarum)"))


```

Set color scales
```{r}
filter_colors_df2 <- data.frame(filter=c("total reads","missing","predator","bait","non-target, non-specific","prey","calibrated"),
                                color=c("black",gray.colors(n=4,start=0.2,end=0.7),"darkred","coral2"))

species_grayscale <- data.frame(common=c("Dungeness crab\n(C. magister)",
                                "Sand shrimp\n (C. franciscorum)","Shiner perch\n(C. aggregata)",
                                "Mud snail\n(B. attramentaria)",
                                "Soft-shell clam\n(M. arenaria)","Staghorn sculpin\n(L. armatus)",
                                "Hairy shore crab\n(H. oregonensis)","Manila clam\n(R. philippinarum)"),
                                color=gray.colors(start=0,end=0.75,n=8))
```




### Read change by step

```{r}
track_reads$step <- factor(track_reads$step,levels=c("total reads","missing","predator","bait","non-target, non-specific","prey","calibrated"))
```


```{r}
plot1 <- ggplot(track_reads,aes(x=step,y=prop_reads_remaining, col=step)) + 
  geom_boxplot()  + 
  geom_jitter(width=0.2) + 
  scale_color_manual(values=filter_colors_df2$color) +
  labs(y="Proportion of reads remaining",x="") + 
  theme_bw() +
  theme(legend.position="none",axis.text.x=element_text(angle=45, hjust=1))
plot1
```



### proportion of all prey represented by calibrated prey

```{r}
plot2 <- track_prey %>% filter(calibrated=="yes") %>%
  pivot_longer(cols=c("pTaxa","pReads"), names_to="metric",values_to="proportion") %>%
  mutate(metric=ifelse(metric=="pTaxa","unique taxa","reads")) %>%
  mutate(site_type=ifelse(site_type=="clam bed","Clam Bed","Slough")) %>%
  ggplot(aes(x=metric,y=proportion)) +
  geom_boxplot(color="coral2") + 
  geom_jitter(alpha=0.6, color="coral2", width=0.2) +
  facet_wrap(~site_type) +
  scale_y_continuous(position="right") +
  labs(x="",y="Proportion calibrated of all prey") +
  theme_bw() + theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(panel.border=element_rect(color="darkred",linewidth=2),
        plot.margin=margin(t=5.5,r=5.5,b=32,l=5.5,unit="pt"))
plot2


# rect <- rectGrob(
#   x = unit(0.05, "in"),
#   y = unit(0.25, "npc") - unit(1, "in"),
#   width = unit(5.35, "in"),
#   height = unit(4.55, "in"),
#   hjust = 0, vjust = 0,
#   gp = gpar(col = "lightblue", fill=NA,lwd=2)
# )
# 
# plot2l <- ggdraw(plot2) + draw_grob(rect)
```






### zoid average crab

Overall
```{r}
plot3 <- fitted_vals_1 %>%
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  left_join(prey_names,by=c("species"="scientific")) %>%
  arrange(mean) %>%
  ggplot(aes(x=reorder(common,-mean),y=mean)) + 
  geom_point() +
  geom_errorbar(aes(ymin=lo,ymax=hi,color=forcats::fct_reorder(common,-mean)), width=0.5) +
  scale_color_grey(name="Prey Species", start=0,end=0.75) +
  labs(x="",y="proportion of DNA in average crab") +
  ylim(c(0.05,0.20)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1), legend.position="none",
        panel.border=element_rect(color="coral2",linewidth=2))
plot3
```


clam bed v slough -- stacked barplot
```{r}
fitted_vals_bysite$site_type <- factor(fitted_vals_bysite$site_type, levels=c("Clam Bed","Slough"))

plot3b <- fitted_vals_bysite %>%
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  dplyr::select(site_type,species,mean) %>% distinct() %>%
  left_join(prey_names,by=c("species"="scientific")) %>%
  left_join(species_grayscale) %>%
  arrange(mean) %>%
  ggplot(aes(x=site_type,y=mean,fill=forcats::fct_reorder(common,-mean),label=common)) + 
  geom_col(position="stack") +
  # geom_text(position="stack", vjust=1,color="white") +
  geom_text(aes(x=site_type,y=mean,label=round(mean,2)), 
            hjust=-1,vjust=1,color="white",position="stack",size=4) +
  geom_hline(aes(yintercept=0)) +
  scale_fill_grey(name="Prey Species", start=0,end=0.75) +
  labs(x="",y="Proportion of DNA in average crab") +
  theme_bw() +
  theme(axis.text.x=element_text(size=12), axis.title.y=element_text(size=12),
        legend.position="right",
        panel.border=element_rect(color="coral2",linewidth=2),
        legend.key.height = unit(2,"line"),
        plot.margin = margin(t=20.5,r=5.5,b=2.5,l=5.5,unit="pt"))
plot3b
```


### Figure 4


```{r}
bottom2 <- plot_grid(plot1 + theme(plot.margin=margin(t=5.5,r=5.5,b=-5,l=5.5,unit="pt")),
                 plot2 + theme(plot.margin=margin(t=5.5,r=5.5,b=27,l=5.5,unit="pt")),
                 ncol=2, labels=c("b","c"),label_y = 1.075)

# plot_grid(plot3b,bottom2,nrow=2,labels=c("","a"),label_y = 1.1, rel_widths=c(1,1))

# png(here('../../','figures','qm_to_zoid_bysite_bar.png'),res=300,width=2000,height=2000)
# plot_grid(top2,plot3b,nrow=2,labels=c("","c"),label_y = 1.1)
# dev.off()

tiff(here('../','figs','Fig4.tif'),res=300,width=2000,height=2200)
plot_grid(plot3b,bottom2,nrow=2,labels=c("","a"),label_y = 2.05, rel_widths=c(1,1), rel_heights=c(1, 0.9))
dev.off()
```


## Supplement



### Fig S6

clam bed v. slough -- means with error
```{r}
fitted_vals_bysite$site_type <- factor(fitted_vals_bysite$site_type, levels=c("Clam Bed","Slough"))
plot3s <- fitted_vals_bysite %>%
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  left_join(prey_names,by=c("species"="scientific")) %>%
  arrange(mean) %>%
  # ggplot(aes(x=reorder(common,-mean),y=mean)) + 
  ggplot(aes(x=forcats::fct_reorder(common,-mean),y=mean)) +
  geom_point(aes(pch=site_type)) +
  geom_errorbar(aes(ymin=lo,ymax=hi,color=forcats::fct_reorder(common,-mean)), width=0.5) +
  facet_wrap(~site_type,scales="free_y", ncol=1,strip.position="right") +
  scale_color_grey(name="Prey Species", start=0,end=0.75) +
  scale_shape_manual(values=c(15,17)) +
  labs(x="",y="Proportion of DNA in average crab") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1), legend.position="none",
        panel.border=element_rect(color="coral2",linewidth=2), 
        panel.spacing = unit(0,"lines"))
plot3s
```

```{r}
png(here('../','figs','S6_Fig.png'),res=300,width=1600,height=1200)
plot3s
dev.off()
```



### Extras 

by site type - point
```{r}
plot_grid(top,plot3s,nrow=2,labels=c("","c"),label_y = 0.15)

png(here('../','figs','FigSy_Fig4-as scatterplot.png'),res=300,width=2000,height=2000)
plot_grid(top,plot3s,nrow=2,labels=c("","c"),label_y = 0.15)
dev.off()
```


### Bonus

across all sites
```{r}
top <- plot_grid(plot1,plot2,ncol=2, labels=c("a","b"),label_y = 0.3)
top
plot_grid(top,plot3,nrow=2,labels=c("","c"),label_y = 0.15)

png(here('../','figs','FigSx_Fig4-all-sites.png'),res=300,width=2000,height=2000)
plot_grid(top,plot3,nrow=2,labels=c("","c"),label_y = 0.15)
dev.off()
```
