---
title: "Shelton et al. Quant eDNA Model"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

**Date written: May 20 2023**

**Author: M Fisher**

**Last Run: `r Sys.Date()`**


## Description

Run the [Shelton et al. (2023)](https://doi.org/10.1002/ecy.3906) quantitative eDNA model to generate calibrate DNA quantities of prey taxa in stomach content samples. 

A note about error messages, from Ryan Kelly:

- Max Tree Depth is a warning you can ignore.  It just means sampling wasn’t as efficient as it could have been, but the results are still totally valid.
- Divergent transitions are intermediate-serious warnings.  These mean that sampling isn’t working quite right, but if there are small numbers of divergent transitions, that’s not too bad.
- Rhat warnings are serious and mean that the model isn’t converging; something is broken.
- Low sample-size warnings aren’t serious; you can either ignore them or else run the chains for longer.


*Run notes:*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(here)
library(ggplot2)
library(ggrepel)
library(beepr)
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

source(here("../","R","calibrate_metabarcoding_RPK.R"))

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```


## Data

This is the data produced in script `qm1_prep_data_for_calibration`. 

```{r}
load(here('../','data','quant_modeling',"QM_dataLIST_sampleRun.rda"))
```

check to make sure sample data loaded in properly
```{r}
unique(EGC_obs[[1]]$Community); unique(EGC_obs[[1]]$species)
unique(EGC_obs[[2]]$station); unique(EGC_obs[[2]]$species)
```

```{r}
n_pcr_samples <- read_csv(here('../','data','quant_modeling','qm_sampleRun_nPCR.csv'))
```


### minimum read depths

wwhat are the read depths going into the quant model?
```{r}
myplot <- EGC_obs[[2]] %>%
  filter(species %in% EGC_obs[[1]]$species) %>%                # species in quant model
  group_by(station,tech) %>% summarise(total_reads=sum(Nreads)) %>%
  ggplot(aes(x=total_reads)) + geom_histogram() +theme_bw()

myplot

png(here('../','data','quant_modeling','quant_modeling_qc','QM_sampleRun_sample-read-depths.png'))
myplot
dev.off()
```

How many crab do we get rid of if we only focus on samples with > 100 reads?
```{r}
reads_min <- 100


crab_low_reads <- EGC_obs[[2]] %>%
  filter(species %in% EGC_obs[[1]]$species) %>%   # filter for the calibrated species
  group_by(station,tech) %>% summarise(total_reads=sum(Nreads)) %>% filter(total_reads < reads_min) %>%
  dplyr::select(station,tech) %>% distinct()

 EGC_obs[[2]] %>%
  filter(species %in% EGC_obs[[1]]$species & station %in% crab_100filter$station) %>% 
   group_by(station,tech) %>% summarise(total_reads=sum(Nreads)) %>% arrange(station,total_reads)
```

Because certain technical replicates have more reads than others, a 100 read filter would only remove 2 crab: WACO21-503 and WACO21-512.

Filter samples according to a minimum read-depth cutoff of n=100
```{r}
EGC_obs[[2]] <- EGC_obs[[2]] %>% anti_join(dplyr::select(crab_low_reads,station,tech))
```

### separate samples by PCR cycle number

The number of indexing PCR cycles varied by sample. When we run the quantitative model, we can only input one PCR cycle number for the sample data. Create single-column data frames for each total PCR cycle value (ater removing the crabs with low read counts identified above)
```{r}
n_pcr_samples <- anti_join(n_pcr_samples %>% mutate(tech=as.character(tech)),
                           crab_low_reads %>% 
                             dplyr::select(station,tech) %>% mutate(tech=as.character(tech)), by=c("sample"="station","tech"))
#format, sample subset had 45 PCR cycles
sub43.samples <- n_pcr_samples %>% filter(PCRcycles==43) %>% dplyr::select(sample) %>% distinct()
sub40.samples <- n_pcr_samples %>% filter(PCRcycles==40) %>% dplyr::select(sample) %>% distinct()
sub47.samples <- n_pcr_samples %>% filter(PCRcycles==47) %>% dplyr::select(sample) %>% distinct()
```


## Run Calibration Model

### nPCR = 40

format 
```{r}
my_dDNA_data <- format_metabarcoding_data(EGC_obs[[2]] %>% filter(station %in% sub40.samples$sample), N_pcr_mock=40,
                                          EGC_obs[[1]],
                                          Level_1_treatment_envir <- "station",
                                          Level_2_treatment_envir <- "tech",
                                          Level_3_treatment_envir <- NA,
                                          Level_1_treatment_mock <- "Community",
                                          Level_2_treatment_mock <- "tech",
                                          Level_3_treatment_mock <- NA)
mydata40 <- makeDesign(my_dDNA_data, N_pcr_cycles = 40)
```


check what species are present
```{r}
mydata40$sp_list
```

run model and save output
```{r}
QM_Bayes_40pcr <- QM_bayes("quant_metabar_multinom.stan", mydata40, NCHAINS = 3, ITER = 5000 )
saveRDS(QM_Bayes_40pcr,file="QM_Bayes_out_40pcr_100rmin.rds")

beepr::beep()
```

*Max tree depth warning, but nothing to worry about.*


What are the alphas?
```{r}
QM_Bayes_40pcr$Bayes_alpha_est
```



### nPCR = 43

format 
```{r}
my_dDNA_data43 <- format_metabarcoding_data(EGC_obs[[2]] %>% filter(station %in% sub43.samples$sample), N_pcr_mock=40,
                                            EGC_obs[[1]],
                                            Level_1_treatment_envir <- "station",
                                            Level_2_treatment_envir <- "tech",
                                            Level_3_treatment_envir <- NA,
                                            Level_1_treatment_mock <- "Community",
                                            Level_2_treatment_mock <- "tech",
                                            Level_3_treatment_mock <- NA)
mydata43 <- makeDesign(my_dDNA_data43, N_pcr_cycles = 43)

```


check what species are present
```{r}
mydata43$sp_list
```

run model and save output
```{r}
QM_Bayes_43pcr <- QM_bayes("quant_metabar_multinom.stan", mydata43, NCHAINS = 3, ITER = 5000 )
saveRDS(QM_Bayes_43pcr,file="QM_Bayes_out_43pcr_100rmin.rds")

beepr::beep()
```

*Max tree depth warning, but nothing to worry about.*


What are the alphas?
```{r}
QM_Bayes_43pcr$Bayes_alpha_est
```




### nPCR = 47

format 
```{r}
my_dDNA_data47 <- format_metabarcoding_data(EGC_obs[[2]] %>% filter(station %in% sub47.samples$sample), N_pcr_mock=40,
                                            EGC_obs[[1]],
                                            Level_1_treatment_envir <- "station",
                                            Level_2_treatment_envir <- "tech",
                                            Level_3_treatment_envir <- NA,
                                            Level_1_treatment_mock <- "Community",
                                            Level_2_treatment_mock <- "tech",
                                            Level_3_treatment_mock <- NA)
mydata47 <- makeDesign(my_dDNA_data47, N_pcr_cycles = 47)
```


check what species are present
```{r}
mydata47$sp_list
```

run model and save output
```{r}
QM_Bayes_47pcr <- QM_bayes("quant_metabar_multinom.stan", mydata47, NCHAINS = 3, ITER = 5000)
saveRDS(QM_Bayes_47pcr,file="QM_Bayes_out_47pcr_100rmin.rds")


beepr::beep()
```



What are the alphas?
```{r}
QM_Bayes_47pcr$Bayes_alpha_est
```



## Combine QM output

```{r}
QM_Bayes_all <- bind_rows((QM_Bayes_40pcr$Bayes_estimates %>%
                             rownames_to_column("crab") %>%
                             pivot_longer(cols=2:(dim(QM_Bayes_40pcr$Bayes_estimates)[2]+1)) %>%
                             mutate(npcr=40)),
                          (QM_Bayes_43pcr$Bayes_estimates %>%
                             rownames_to_column("crab") %>%
                             pivot_longer(cols=2:(dim(QM_Bayes_43pcr$Bayes_estimates)[2]+1)) %>%
                             mutate(npcr=43)),
                          (QM_Bayes_47pcr$Bayes_estimates %>%
                             rownames_to_column("crab") %>%
                             pivot_longer(cols=2:(dim(QM_Bayes_47pcr$Bayes_estimates)[2]+1))) %>%
                            mutate(npcr=47))
bayes_clean <- filter(QM_Bayes_all, value > 0.001) %>%  mutate(name=ifelse(name=="Mya","Mya arenaria",name))
```


```{r}
bayes_clean %>% ggplot(aes(x=crab,y=value,fill=name)) +
  geom_col() + 
  theme(axis.text.x=element_blank())
```

save 
```{r}
write_csv(bayes_clean, here('../','data','quant_modeling','QM_BayesOut_sampleRun_means_filtered.csv'))
```


### save for fitting zoid!

```{r}
all_dDNA_data_Observation <- my_dDNA_data$Observation %>% mutate(npcr=40) %>%
  bind_rows(my_dDNA_data43$Observation %>% mutate(npcr=43)) %>%
  bind_rows(my_dDNA_data47$Observation %>% mutate(npcr=47))

saveRDS(all_dDNA_data_Observation,here('data','quant_modeling','QM_BayesOut_sampleRun_dDNAobservations.rds'))
```
