---
title: "Run Zoid - Average Crab"
author: "M Fisher"
date: "2023-08-22"
output: html_document
---

# Description

Run a hierarchical model to estimate the diet of the “average crab,” using a zero-inflated dirichlet distribution (species *i* crab *j* represents a draw from a beta distribution $p_{ij}$).
 
 All means need to sum to 1 in the average crab; use a dirichlet distribution

 Start with (a) Bayes est from QM model, and (b) total read count fed into the QM model,
    to calculate the number of expected reads per species (using a multinomial draw).
    Then create matrix of expected reads per species (columns) per crab (rows)
     and feed into zoid.
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(zoid)
library(tidyverse)
library(here)
library(magrittr)
library(forcats)
library(cowplot)

knitting=TRUE
```


## Data

QM model output, run by pcr cycles. Only includes the posterior mean estimates that match to species actually present in the diet data. 
```{r}
QM_Bayes_filter <- read_csv(here('../','data','quant_modeling','QM_BayesOut_sampleRun_means_filtered.csv'))

observed_prey_vec <- unique(filter(QM_Bayes_filter, name!="Carcinus maenas") %>% pull(name))
```

the diet data (sequencing reads for modeled species) that was used to run the QM
```{r}
all_dDNA_data_Observation <- readRDS(here('../','data','quant_modeling','QM_BayesOut_sampleRun_dDNAobservations.rds'))
```

get the average total reads per crab, across technical replicates
```{r}
dDNA_avgReads <- all_dDNA_data_Observation %>%
  group_by(Sample,level1,level2,npcr) %>%
  summarize(sample_reads=sum(Nreads)) %>%
  ungroup() %>%
  group_by(level1,npcr) %>% nest() %>%
  mutate(mean_reads= map(data, ~ mean(x=.x$sample_reads))) %>%
  mutate(techs = map(data, ~ dim(.x)[1])) %>%
  dplyr::select(level1,techs,npcr, mean_reads) %>% 
  unnest(cols=c(level1,techs,npcr,mean_reads))
```

read in site information
```{r}
metadat <- read_csv(here('../','data','metadata','Willapa Bay EGC Samples - Sample Data.csv'))
metadat %<>% dplyr::select(Sample_label,Site_Name) %>%
  mutate(site=ifelse(grepl("Nahcotta",Site_Name),"Nahcotta",
                     ifelse(grepl("Stackpole",Site_Name),"Stackpole",
                            ifelse(grepl("Oysterville",Site_Name),"Oysterville",
                                   ifelse(grepl("Long Beach",Site_Name),"Long Beach",NA))))) %>%
  dplyr::select(-Site_Name)
```

## Format Input

**if 'zoid_input_Bayes_est_reads.rds' has already been created, skip this!**

### clean up QM Bayes estimates

some samples were sequenced twice: 
```{r echo=FALSE, eval=TRUE}
#check proportions
QM_Bayes_filter %>% ggplot(aes(x=crab,y=value,fill=name)) +
  geom_col() + 
  theme(axis.text.x=element_blank())
```
how different are the results for those samples?
```{r, eval=TRUE}
fix.sample <- QM_Bayes_filter %>% group_by(crab) %>% mutate(p=sum(value)) %>% filter(p>1) %>%
  dplyr::select(-sum_reads) %>%
  pivot_wider(names_from=npcr,values_from=value)

fix.sample
```

that's cool, not that different between runs. take the posterior means for the sample run with the *least* PCR cycles. just use those. 
```{r eval=TRUE}
fixed.sample <- QM_Bayes_filter %>%
  filter(crab %in% fix.sample$crab) %>%
  group_by(crab,name) %>% arrange(npcr) %>% mutate(ranks = order(order(npcr))) %>%
  filter(ranks==1) %>%
  dplyr::select(-ranks)

bayes_clean <- QM_Bayes_filter %>% filter(!(crab %in% fixed.sample$crab)) %>%
  bind_rows(dplyr::select(fixed.sample, all_of(colnames(QM_Bayes_filter))))
```


```{r echo=FALSE, eval=TRUE}
#check proportions again!
bayes_clean %>% ggplot(aes(x=crab,y=value,fill=name)) +
  geom_col() + 
  theme(axis.text.x=element_blank())
```


### Combine Bayes estimates with reads data

add total sample read depths to the Bayes estimates
```{r eval=TRUE}
bayes_clean_reads <- bayes_clean %>% dplyr::select(-sum_reads) %>%
  left_join(dplyr::select(dDNA_avgReads, level1,npcr,mean_reads) %>% distinct(),
            by=c("crab"="level1","npcr"))
sum(is.na(bayes_clean_reads$mean_reads))
```

calculate reads per species by drawing from a multinomially distributed random number vector
```{r eval=TRUE}
set.seed(555)
zoid_input <- bayes_clean_reads %>%
  group_by(crab) %>%
  nest() %>%
  summarise(new_counts=map(data, ~rmultinom(n=1,size=unique(.x$mean_reads), prob=.x$value))) %>%
  unnest_longer(new_counts, indices_include=FALSE) %>% arrange(crab,desc(new_counts)) %>%
  bind_cols(bayes_clean_reads %>% rename(crab2=crab) %>% arrange(crab2,desc(value))) %>%
  mutate(sp_reads=new_counts[,1]) %>% dplyr::select(-new_counts,-crab2)
```

Because many species contribute super low (but non-zero) proportion of DNA, the `rmultinom` function may return 0 reads. re-run `rmultinom` with a new seed until there are no zeroes remaining.
```{r eval=TRUE}
s <- 560
while(any(zoid_input$sp_reads==0)){
  set.seed(s)
  zoid_input %<>% mutate(sp_reads=ifelse(sp_reads==0, rmultinom(n=1,size=mean_reads, prob=value),sp_reads))
  s <- s+5
  message(sum(zoid_input$sp_reads==0), " zeros remaining.")
}
```

plot the QM output data
```{r eval=TRUE}
if(!knitting){
  
png(here('../','data','quant_modeling','quant_modeling_qc','zoid_input_Bayes_est.png'),res=300,height=1500,width=1500)
zoid_input %>% mutate(name=ifelse(name=="Mya","Mya arenaria",name)) %>%
  pivot_longer(cols=c("value","sp_reads"),names_to="group",values_to="value") %>%
  mutate(group=ifelse(group=="value","propDNA",
                      ifelse(group=="sp_reads","readCount_sampled",NA))) %>%
  ggplot(aes(x=name,y=value)) +
  geom_point(aes(col=name)) + geom_violin(alpha=0.5, aes(col=name), fill=NA) +
  facet_wrap(~group,scales="free_y") +
  labs(x="",y="proportion DNA / read counts") +
  theme_bw() + theme(axis.text.x=element_text(angle=45, hjust=1),
                     legend.position="none",
                     plot.margin=unit(c(top=5.5,right=5.5,bottom=5.5,left=20.5),"points"))
dev.off()

}
```

make matrix of the expected number of reads
```{r eval=TRUE}
bayes_wide <- zoid_input %>% 
  dplyr::select(crab,npcr,name,sp_reads) %>%
  pivot_wider(names_from="name",values_from="sp_reads", values_fill=0)
```

### Attach site info

```{r eval=TRUE}
bayes_wide %<>% left_join(metadat,by=c("crab"="Sample_label"))
bayes_wide %<>% mutate(site_type=ifelse(site %in% c("Nahcotta","Stackpole"),"slough","clam bed"))
```

### SAVE the ZOID INPUT DATAFRAME
```{r}
if(!knitting){
  
saveRDS(bayes_wide,here('../','data','quant_modeling','zoid_input_Bayes_est_rmultinomReads.rds'))
  
}
```


### Remove predator 

Remove the predator information from the data frame
```{r}
bayes_wide_prey <- dplyr::select(bayes_wide, -`Carcinus maenas`)
```


## Zoid : prey all sites

### run zoid

create the design matrix (site / site type information per observation, aka per crab) and the data matrix (read counts only)
```{r}
design_matrix = bayes_wide_prey[,names(bayes_wide_prey)%in%c("site","site_type")==TRUE]
data_matrix = bayes_wide_prey[,names(bayes_wide_prey)%in%c("site","site_type","crab","npcr")==FALSE]
design_matrix$y = 1 # dummy variable
```

run zoid!
```{r}
fit_1_prey <- fit_zoid(data_matrix = as.matrix(data_matrix/100),
                       overdispersion = TRUE,
                       chains=4,
                       iter=10000)
```

### format / explore output
```{r}
fitted_vals_1 = get_fitted(fit_1_prey)
fitted_vals_1 %<>% left_join(data.frame(group=seq(1,9,by=1),
                                         species=c("Crangon franciscorum","Hemigrapsus oregonensis","Leptocottus armatus",
                                                   "Mya","Metacarcinus magister","Cymatogaster aggregata","Batillaria attramentaria",
                                                   "Ruditapes philippinarum","other")), by=c("group"))
```

```{r echo=FALSE,fig.width=8,fig.height=6}
avg_crab1 <- fitted_vals_1 %>%
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  arrange(mean) %>%
  ggplot(aes(x=reorder(species,-mean),y=mean)) + 
  geom_point() +
  geom_errorbar(aes(ymin=lo,ymax=hi,color=forcats::fct_reorder(species,-mean)), width=0.5) +
  scale_color_grey(name="Prey Species", start=0,end=0.75) +
  labs(x="",y="proportion of DNA") +
  ylim(c(0.05,0.20)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1), legend.position="none")

avg_crab2 <- fitted_vals_1 %>% 
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  mutate(avg_crab=1) %>%
  dplyr::select(avg_crab,species,mean) %>% distinct() %>%
  ggplot(aes(x=avg_crab,y=mean, fill=forcats::fct_reorder(species,-mean))) + 
  geom_col(col="grey30") +
  geom_text(aes(x=avg_crab,y=mean,label=round(mean,2)), hjust=-1,vjust=1,color="white",position="stack") +
  labs(x="",y="mean proportion of DNA") +
  scale_fill_grey(name="Prey Species", start=0,end=0.75) +
  theme_bw() +
  theme(axis.text.x=element_blank(),plot.margin=margin(5.5,5.5,75,5.5,unit="pt"),
        panel.grid.major.x = element_blank(),axis.ticks.x=element_blank(),panel.grid.minor.x=element_blank())
plot_grid(avg_crab1,avg_crab2, rel_widths=c(0.9,1))
```

```{r}
png(here('zoid_fittedOut_rmultinomReads.png'), res=300,height=1500,width=1500)
avg_crab1 + theme(plot.margin=margin(5.5,5.5,5.5,20, unit="pt"))
dev.off()
```


### save output

Save fitted vals! use these later to make the combined figure 
```{r}
write_csv(fitted_vals_1, here('../','data','quant_modeling','zoid_fittedOut_preyD100.csv'))
```



## Zoid : by site type

### prep data
get two matrices, one for each site
```{r}
bayes_prey_slough <- filter(bayes_wide_prey, site_type=="slough")
bayes_prey_clam   <- filter(bayes_wide_prey, site_type=="clam bed")
```

remove empty columns
```{r}
to_remove <- observed_prey_vec[which(colSums(bayes_prey_clam[,observed_prey_vec])==0)]
clambed_observed_prey <- observed_prey_vec[which(colSums(bayes_prey_clam[,observed_prey_vec])>0)]
if(length(to_remove) > 0){
  bayes_prey_clam %<>% dplyr::select(-all_of(to_remove))
  colSums(bayes_prey_clam[,clambed_observed_prey])
}
```

```{r}
to_remove <- observed_prey_vec[which(colSums(bayes_prey_slough[,observed_prey_vec])==0)]
slough_observed_prey <- observed_prey_vec[which(colSums(bayes_prey_slough[,observed_prey_vec])>0)]
if(length(to_remove) > 0){
  bayes_prey_slough %<>% dplyr::select(-all_of(to_remove))
  colSums(bayes_prey_slough[,slough_observed_prey])
}
```

### zoid for clam beds

#### run zoid
set up matrices
```{r}
data_matrix = bayes_prey_clam[,names(bayes_prey_clam)%in%c("site","site_type","crab","npcr")==FALSE]
design_matrix$y = 1 # dummy variable
```


zoid still needs the data matrix to be divided by 100 (doesn't run with only data matrix, or divided by 10)
```{r}
set.seed(555)
fit.clambed <- fit_zoid(data_matrix = as.matrix(data_matrix/100),
                    overdispersion = TRUE,
                    chains=3,
                    iter=10000)
```


#### format / explore output
explore posterior means
```{r}
fitted_vals_clambed = get_fitted(fit.clambed)
fitted_vals_clambed %<>% left_join(data.frame(group=seq(1,8,by=1),
                                         species=colnames(data_matrix)), by=c("group"))
```

```{r echo=FALSE,fig.width=8,fig.height=6}
avg_crab1 <- fitted_vals_clambed %>%
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  arrange(mean) %>%
  ggplot(aes(x=reorder(species,-mean),y=mean)) + 
  geom_point() +
  geom_errorbar(aes(ymin=lo,ymax=hi,color=forcats::fct_reorder(species,-mean)), width=0.5) +
  scale_color_grey(name="Prey Species", start=0,end=0.75) +
  labs(x="",y="proportion of DNA") +
  ylim(c(0.05,0.20)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1), legend.position="none")

avg_crab2 <- fitted_vals_clambed %>% 
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  mutate(avg_crab=1) %>%
  dplyr::select(avg_crab,species,mean) %>% distinct() %>%
  ggplot(aes(x=avg_crab,y=mean, fill=forcats::fct_reorder(species,-mean))) + 
  geom_col(col="grey30") +
  geom_text(aes(x=avg_crab,y=mean,label=round(mean,2)), hjust=-1,vjust=1,color="white",position="stack") +
  labs(x="",y="mean proportion of DNA") +
  scale_fill_grey(name="Prey Species", start=0,end=0.75) +
  theme_bw() +
  theme(axis.text.x=element_blank(),plot.margin=margin(5.5,5.5,75,5.5,unit="pt"),
        panel.grid.major.x = element_blank(),axis.ticks.x=element_blank(),panel.grid.minor.x=element_blank())
plot_grid(avg_crab1,avg_crab2, rel_widths=c(0.9,1))
```

```{r}
png(here('zoid_fittedOut_CLAMBED_rmultinomReads.png'), res=300,height=1500,width=1500)
avg_crab1 + theme(plot.margin=margin(5.5,5.5,5.5,20, unit="pt"))
dev.off()
```


#### save
Save fitted vals! use these later to make the combined figure 
```{r}
write_csv(fitted_vals_clambed, here('../../','data','zoid','zoid_fittedOut_CLAMBED_preyD100.csv'))
```


### zoid for slough

#### run zoid
set up matrices
```{r}
data_matrix = bayes_prey_slough[,names(bayes_prey_slough)%in%c("site","site_type","crab","npcr")==FALSE]
design_matrix$y = 1 # dummy variable
```

zoid still needs the data matrix to be divided by 100 (doesn't run with only data matrix, or divided by 10, or with prior_sd = 1,2)

```{r}
set.seed(555)
fit.slough <- fit_zoid(data_matrix = as.matrix(data_matrix/100),
                    overdispersion = TRUE,
                    chains=3,
                    iter=10000)
```

```{r}
fitted_vals_slough = get_fitted(fit.slough)
fitted_vals_slough %<>% left_join(data.frame(group=1:dim(data_matrix)[2],
                                         species=colnames(data_matrix)), by=c("group"))
```


#### format / explore output

explore fitted vals
```{r echo=FALSE,fig.width=8,fig.height=6}
avg_crab1 <- fitted_vals_slough %>%
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  arrange(mean) %>%
  ggplot(aes(x=reorder(species,-mean),y=mean)) + 
  geom_point() +
  geom_errorbar(aes(ymin=lo,ymax=hi,color=forcats::fct_reorder(species,-mean)), width=0.5) +
  scale_color_grey(name="Prey Species", start=0,end=0.75) +
  labs(x="",y="proportion of DNA") +
  ylim(c(0,0.60)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1), legend.position="none")

avg_crab2 <- fitted_vals_slough %>% 
  mutate(species=ifelse(species=="Mya","Mya arenaria",species)) %>%
  mutate(avg_crab=1) %>%
  dplyr::select(avg_crab,species,mean) %>% distinct() %>%
  ggplot(aes(x=avg_crab,y=mean, fill=forcats::fct_reorder(species,-mean))) + 
  geom_col(col="grey30") +
  geom_text(aes(x=avg_crab,y=mean,label=round(mean,2)), hjust=-1,vjust=1,color="white",position="stack") +
  labs(x="",y="mean proportion of DNA") +
  scale_fill_grey(name="Prey Species", start=0,end=0.75) +
  theme_bw() +
  theme(axis.text.x=element_blank(),plot.margin=margin(5.5,5.5,75,5.5,unit="pt"),
        panel.grid.major.x = element_blank(),axis.ticks.x=element_blank(),panel.grid.minor.x=element_blank())
plot_grid(avg_crab1,avg_crab2, rel_widths=c(0.9,1))
```

```{r}
png(here('zoid_fittedOut_CLAMBED_rmultinomReads.png'), res=300,height=1500,width=1500)
avg_crab1 + theme(plot.margin=margin(5.5,5.5,5.5,20, unit="pt"))
dev.off()
```

#### save

Save fitted vals! use these later to make the combined figure 
```{r}
write_csv(fitted_vals_slough, here('../../','data','zoid','zoid_fittedOut_SLOUGH_preyD100.csv'))
```



# Save final zoid

save all final zoid runs to a list
Save fitted vals! use these later to make the combined figure 
```{r}
zoid.list <- list(fit_1_prey,fit.clambed,fit.slough)
names(zoid.list) <- c("all crab","clam bed","slough")
saveRDS(zoid.list, here('../../','data','zoid','zoid_fittedOut_preyD100.rds'))
```
