---
title: "Results: Prey Diversity"
author: "M Fisher"
date: "2023-03-26"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Description

Summarize the following across sequencing runs:

1. Sample sizes:
  a. how many crab were sequenced?
  b. how many crab had prey information?

2. Basic prey counts:
  a. taxonomic depth - what proportion of reads were identified down to genus or species? 
  b. taxonomic breadth - What was the unique number of taxa identified per site/month? per site?
  
3. Per site prey information:
  a. for each taxon, what proportion of crabs at each site/month had that prey item? *(qc sidebar: of those, what proportion of technical replicates had that prey item?)*
  b. eDNA index between sites


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(janitor)
library(magrittr)
library(vegan)
library(stargazer)

# library(devtools)
# install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(pairwiseAdonis)

# devtools::install_github("jfq3/ggordiplots")
library(ggordiplots)
# source(here('R','gg_ordiplot_custom.R'))
source(here('R','eDNA_index.R'))
source(here('R','get_unique_taxa.R'))

##### User inputs #####
run.nums <- c(2,3,4,5)
marker  <- 'BF3'
meta_dir <- 'data/metadata'
results_dir <- 'data/results'
# dir2 <- '../Green-crab-dDNA'
```
```{r data}
# metadata from trapping
meta.trap <- read_csv(here(meta_dir,'Willapa Bay EGC Samples - Sample Data.csv'))

# metadata for run
meta.run <- read_csv(here(meta_dir,'runs2-5_BF3_run_metadata.csv'))

# sequencing data file - final manual filtered
dat  <- read_csv(here(results_dir, 'allRuns_BF3_filtered_FINAL_unique_taxa.csv'))

# count of crab dissected by site
site_coords <- read_csv(here(meta_dir,'site_coords.csv'))

set.seed(555)
```

# Sample Size

How many crab were sequenced, per site and month?

```{r}
meta.run %>%
  filter(!is.na(site_month)) %>%
  group_by(site_month) %>%
  summarise(`sequenced`=length(unique(sample)))

```


## n crab per step

```{r fig.height=3, fig.width=5}
dissected <- site_coords %>% group_by(site_name) %>% summarise(n_dissected=sum(n_dissected)) %>%
  rename(Site=site_name)

sequenced <- meta.run %>%
  filter(!is.na(site_month)) %>%
  separate(site_month, into=c("Site","Month"), sep="-") %>%
  group_by(Site) %>%
  summarise(`sequenced`=length(unique(sample)))

success <- dat %>%
  filter(!is.na(site_month)) %>%
  separate(site_month, into=c("Site","Month"), sep="-") %>%
  group_by(Site) %>%
  summarise(`with prey`=length(unique(sample))) %>%
  mutate(`with prey`=ifelse(is.na(`with prey`),0,`with prey`))

ncrab_bySite <- full_join(sequenced,success,by="Site") %>%
  mutate(`seq with prey` = round(`with prey`/`sequenced`,2)) 
ncrab_bySite
```


```{r}
dissectedM <- site_coords %>% group_by(sampling_month) %>% summarise(n_dissected=sum(n_dissected)) %>%
  rename(Month=sampling_month) %>%
  mutate(Month=ifelse(Month=="July","Jul",
                      ifelse(Month=="September","Sep",Month)))
sequencedM <- meta.run %>%
  filter(!is.na(site_month)) %>%
  separate(site_month, into=c("Site","Month"), sep="-") %>%
  group_by(Month) %>%
  summarise(`sequenced`=length(unique(sample)))

successM <- dat %>%
  filter(!is.na(site_month)) %>%
  separate(site_month, into=c("Site","Month"), sep="-") %>%
  group_by(Month) %>%
  summarise(`with prey`=length(unique(sample))) %>%
  mutate(`with prey`=ifelse(is.na(`with prey`),0,`with prey`))

ncrab_byMonth <- full_join(sequencedM,successM,by="Month") %>%
  mutate(`seq with prey` = round(`with prey`/`sequenced`,2)) 
ncrab_byMonth
```
<br>

## carapace widths

Mean carapace widths (mm): 
```{r}
cw <- meta.trap %>%
  mutate(sequenced=ifelse(Sample_label %in% meta.run$sample, "Sequenced", "Trapped"))

cw %<>% mutate(site=ifelse(grepl("Stackpole",Site_Name), "Stackpole",
                                 ifelse(grepl("Oysterville",Site_Name), "Oysterville",
                                        ifelse(grepl("Nahcotta",Site_Name),"Nahcotta",
                                               ifelse(grepl("Long Beach",Site_Name),"Long Beach", NA)))))

cw %>% group_by(site,sequenced) %>% summarise(mean_cw=mean(CW_mm)) %>% pivot_wider(names_from=sequenced,values_from=mean_cw)
```

Range for **collected** crab?
```{r}
cw %>% group_by(Sex) %>% summarise(n.crab=length(unique(Sample_label)),
                                                                      min.cw=min(CW_mm, na.rm = TRUE),
                                                                      max.cw=max(CW_mm, na.rm = TRUE))
```

Range for **sequenced** crab?
```{r}
cw %>% filter(sequenced=="Sequenced") %>% group_by(Sex) %>% summarise(n.crab=length(unique(Sample_label)),
                                                                      min.cw=min(CW_mm, na.rm = TRUE),
                                                                      max.cw=max(CW_mm, na.rm = TRUE))
```

Range for sequenced crab **with prey**?
```{r}
meta.trap %>%
  mutate(with_prey=ifelse(Sample_label %in% dat$sample, "with Prey", "Trapped")) %>%
  mutate(site=ifelse(grepl("Stackpole",Site_Name), "Stackpole",
                                 ifelse(grepl("Oysterville",Site_Name), "Oysterville",
                                        ifelse(grepl("Nahcotta",Site_Name),"Nahcotta",
                                               ifelse(grepl("Long Beach",Site_Name),"Long Beach", NA))))) %>% 
  filter(with_prey=="with Prey") %>% group_by(Sex) %>% summarise(n.crab=length(unique(Sample_label)),
                                                                      min.cw=min(CW_mm, na.rm = TRUE),
                                                                      max.cw=max(CW_mm, na.rm = TRUE))
```





# Alpha Diversity 

What was the unique number of taxa identified?

## Across sites

All sites
```{r}
length(unique(get_unique_taxa(taxa.df=dat, level="all", return.removed=FALSE) %>%
  pull(taxon)))
```

By site type
```{r}
get_unique_taxa(taxa.df = dat %>%  separate(site_month, into=c("site","month"), sep="-"),
                level="site",return.removed=FALSE) %>%
  group_by(site) %>% summarise(n=length(unique(taxon)))
```


Take a quick look at what was removed
```{r eval=FALSE}
View((get_unique_taxa(taxa.df=dat %>%  separate(site_month, into=c("site","month"), sep="-"), level="site", return.removed=TRUE))[[2]])
```


## Per crab stomach


```{r fig.height=3, fig.width=4}
crab.tax.breadth <-dat %>% 
  filter(!is.na(site_month)) %>%
  group_by(site_month, sample) %>% 
  summarise(n_taxa=length(unique(species)))

ggplot(crab.tax.breadth, aes(x=n_taxa)) +
  geom_histogram() + 
  labs(x="Number of Prey Taxa in Stomach", y="Number of Crabs") +
  scale_x_continuous(breaks=seq(1,13,by=2)) +
  theme_bw()
```
```{r eval=FALSE}
png(here('results','figures','alpha-diversity_per-crab_CEGblast.png'),res=300,height=800,width=800)
ggplot(crab.tax.breadth, aes(x=n_taxa)) +
  geom_histogram() + 
  labs(x="Number of Prey Taxa in Stomach", y="Number of Crabs") +
  scale_x_continuous(breaks=seq(1,13,by=2)) +
  theme_bw()
dev.off()
```
<br>

What is in the crab with > 10 species??
```{r}
crab <- (filter(crab.tax.breadth, n_taxa > 10))$sample
crab
dat %>% filter(sample==crab) %>%
  group_by(class,taxon,MiSeqRun) %>% summarise(totalReads=sum(nReads), n_tech=length(unique(tech)))
```
<br>

How does number of prey taxa in stomach compare to read depth from that crab?

How does number of prey taxa in stomach compare to fullness for that crab?



# Beta diversity

What is the $\beta$ diversity for prey taxon? (the ratio between regional and local prey diversity) The greater the similarity in community composition between multiple communities, the lower the value of β-diversity for that set of communities.
```{r}
prey.site.mat <- dat %>%
  separate(site_month,into=c("site","month"),sep="-",remove=FALSE) %>%
  dplyr::select(site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="site", names_from="taxon",values_from="presence", values_fill=0)

bdiv <- as.matrix(betadiver(prey.sp.site.mat, method="w")) 
colnames(bdiv) <-c(prey.sp.site.mat$site); rownames(bdiv) <- c(prey.sp.site.mat$site)
bdiv
```


How much changes if we just look at species-level identifications?
```{r}
preySP.site.mat <- dat %>%
  filter(rank=="species") %>%
  separate(site_month,into=c("site","month"),sep="-",remove=FALSE) %>%
  dplyr::select(site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="site", names_from="taxon",values_from="presence", values_fill=0)

bdivSP <- as.matrix(betadiver(preySP.site.mat, method="w")) 
colnames(bdivSP) <-c(preySP.site.mat$site); rownames(bdivSP) <- c(preySP.site.mat$site)
bdivSP
```


Oysterville v Long Beach beta diversity is reduced (by 0.04); as is Oysterville and Long Beach v. Nahcotta (by 0.02, 0.06 respectively).




# Specificity

what **percent** of reads in the final data set were identified down to genus or species for prey species? 

Across all sites
```{r}
total.reads <- sum(dat$nReads)
all.tax.depth <- dat %>%
  group_by(rank) %>%
  summarise(level_reads=sum(nReads),
            prop_level_reads=level_reads/total.reads,
            per_level_reads=prop_level_reads*100)
printvec <- round(pull(all.tax.depth,per_level_reads),3)
names(printvec)<- pull(all.tax.depth,rank)

printvec[printvec %>% order]
```
<br> 

At each site
```{r}
site.tax.depth <- dat %>%
  separate(site_month,into=c("site","month"),sep="-",remove=FALSE) %>%
  group_by(site) %>% mutate(site_reads=sum(nReads)) %>%
  ungroup() %>%
  group_by(site,rank) %>%
  summarise(level_reads=sum(nReads),
            prop_level_reads=level_reads/site_reads,
            per_level_reads=round(prop_level_reads*100,3)) %>%
  ungroup() %>% distinct()

site.tax.depth %>%
  dplyr::select(site,rank,per_level_reads) %>%
  pivot_wider(names_from=rank,values_from=per_level_reads)
```


