---
title: "Differences between sites"
author: "M Fisher"
date: "2023-05-26"
output: 
  html_document:
    toc: yes
---

# Description

Explore impacts of using different distance metrics on significance of between-site differences. The following are run using (1) presence / absence data and the Jaccard dissimilarity index, and (2) the semi-quantitative eDNA index and the Bray-Curtis distance.

- PERMANOVA

- PERMDISP + PCoA visualization

- Post-hoc permutation tests on pairs of sites. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(janitor)
library(magrittr)
library(vegan)
library(stargazer)

# library(devtools)
# install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(pairwiseAdonis)

# devtools::install_github("jfq3/ggordiplots")
library(ggordiplots)
source(here('R','gg_ordiplot_custom.R'))
source(here('R','eDNA_index.R'))

# User inputs
indir   <- 'results'
run.nums <- c(2,3,4,5)
marker  <- 'BF3'
metadata_dir <- 'data/metadata'
writing <- FALSE
```
```{r data, include=FALSE}
# metadata for run
meta.run <- read_csv(here('data','metadata','runs2-5_BF3_run_metadata.csv'))

dat <- read_csv(here('data','results','allRuns_BF3_filtered_FINAL_unique_taxa.csv'))

site_coords <- read_csv(here('data','metadata','site_coords.csv'))
```


# Presence

## PERMANOVA / PERMDISP

First, run a PERMANOVA on a presence/absence matrix, using the Jaccard dissimilarity index.
```{r}
crab.sp.mat <- dat %>%
  dplyr::select(sample, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="sample", names_from="taxon",values_from="presence", values_fill=0)
crabs <- crab.sp.mat$sample
crab.sp.mat <- as.matrix(crab.sp.mat %>% dplyr::select(-sample))
rownames(crab.sp.mat) <- crabs

indat <- dat %>%
  separate(site_month,into=c("site","month"),sep="-",remove=FALSE) %>%
  dplyr::select(sample, site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols=c("site","sample"), names_from="taxon",values_from="presence", values_fill=0)

```

```{r echo=TRUE}
site.perm <- adonis2(indat[,3:dim(indat)[2]]~site, data=indat, method="jaccard")
site.perm
```

The permanova is significant. 

Is the significance due to the differences in centroids or due to the differences in dispersion of samples in principal coordinate space of dissimilarity (PCoA)?
```{r}
indat.dist <- vegdist(x=indat[,3:dim(indat)[2]],method="jaccard")
site.perm.betadisper <- betadisper(d=indat.dist, group=indat$site)
plot(site.perm.betadisper)
anova(betadisper(d=indat.dist, group=indat$site))
adonis2(dist(site.perm.betadisper$distances)~site.perm.betadisper$group)
```
Borderline but *technically* non-significant result from PERMDISP: groups do not significantly differ in dispersion and therefore the differences are due to differences in location 


PCoA from the PERMDISP

```{r}
indat.sites <- indat %>% mutate(site_type=ifelse(site %in% c("Oysterville","Long Beach"), "Clam Bed", "Slough"))
myplotPRESENCE <- gg_ordiplot_egc(ord=site.perm.betadisper, groups=indat$site, hull=FALSE, ellipse=FALSE, label=FALSE, spiders=TRUE)

myplotPRESENCE$plot + 
  scale_shape_manual(values=c(15,17), name="Site Type") +
  scale_linetype_manual(values=c(1,1), name="Site Type")  + 
  scale_color_manual(values=c("#018571","#f8ab5b","#4fc1af","#a6611a"), name="Sites") +
  labs(x="PCoA Dim 1",y="PCoA Dim 2") +
  theme_bw() + theme(axis.text=element_blank(),axis.ticks=element_blank())
```


## Post-hoc pairwise tests

So, which paired location comparisons that were significantly different?
```{r echo=TRUE}
pairwise.adonis2(indat.dist~site,data=indat,nperm=999)
```



# eDNA Index

eDNA index based on read counts that are calculated by: (a) summing reads across all replicates, and then (b) dividing by the number of replicates for that crab, because different crab have different numbers of replicates, inflating read values.
```{r}
crab.techs <- dat %>%
  group_by(sample) %>% summarise(nTech=length(unique(tech)))
crab.reads.mat <- dat %>%
  group_by(sample, taxon) %>% summarise(techReads=sum(nReads)) %>%
  left_join(crab.techs) %>%
  mutate(adjReads=techReads/nTech) %>%
  dplyr::select(sample,taxon,adjReads) %>%
  pivot_wider(id_cols="sample", names_from="taxon",values_from="adjReads", values_fill=0) %>%
  column_to_rownames("sample")

index.mat <- eDNAINDEX(x=t(crab.reads.mat))

index.df <- as.data.frame(t(index.mat)) %>%
  rownames_to_column("sample") %>% 
  pivot_longer(cols=2:(dim(index.mat)[1]+1), names_to="taxon",values_to="DNAindex") %>%
  pivot_wider(names_from=taxon,values_from=DNAindex)
```
```{r}
index2.indat <- left_join(index.df, indat %>% dplyr::select(site,sample) %>% distinct(),by="sample") %>%
  dplyr::select(site,all_of(colnames(index.df)))
```

## PERMANOVA / PERMDISP

Bray-curtis
```{r echo=TRUE}
site.index2.perm <- adonis2(index2.indat[,3:dim(index2.indat)[2]]~site, data=index2.indat, method="bray")
site.index2.perm
```
The permanova is significant

Is the significance due to the differences in centroids or due to the differences in dispersion of samples in principal coordinate space of dissimilarity (PCoA)?
```{r}
index2.indat.dist <- vegdist(x=index2.indat[,3:dim(index2.indat)[2]],method="bray")
site.index2.perm.betadisper <- betadisper(d=index2.indat.dist, group=index2.indat$site)
plot(site.index2.perm.betadisper)
anova(betadisper(d=index2.indat.dist, group=index2.indat$site))
adonis2(dist(site.index2.perm.betadisper$distances)~site.index2.perm.betadisper$group)
```
Non-significant result from PERMDISP: groups do not differ in dispersion and therefore the differences are entirely due to differences in location 


PCoA from the PERMDISP

```{r}
myplotINDEX <- gg_ordiplot_egc(ord=site.index2.perm.betadisper, groups=index2.indat$site, hull=FALSE, ellipse=FALSE, label=FALSE, spiders=TRUE)

myplotINDEX$plot + 
  scale_shape_manual(values=c(15,17), name="Site Type") +
  scale_linetype_manual(values=c(1,1), name="Site Type")  + 
  scale_color_manual(values=c("#018571","#f8ab5b","#4fc1af","#a6611a"), name="Sites") +
  labs(x="PCoA Dim 1",y="PCoA Dim 2") +
  theme_bw() + theme(axis.text=element_blank(),axis.ticks=element_blank(),legend.text=element_text(size=14),
                     legend.title=element_text(size=14))
```

## Post-hoc pairwise tests

So, which paired location comparisons that were significantly different?
```{r echo=TRUE}
pairwise.adonis2(index2.indat.dist~site,data=index2.indat,nperm=999)
```


# Figure 3
```{r}

fig3a <- myplotPRESENCE$plot + 
  scale_shape_manual(values=c(15,17), name="Site Type") +
  scale_linetype_manual(values=c(1,1), name="Site Type")  + 
  scale_color_manual(values=c("#018571","#f8ab5b","#4fc1af","#a6611a"), name="Sites") +
  labs(x="PCoA Dim 1",y="PCoA Dim 2") +
  theme_bw() + theme(axis.text=element_blank(),axis.ticks=element_blank(), legend.position="none")

fig3b <- myplotINDEX$plot + 
  scale_shape_manual(values=c(15,17), name="Site Type") +
  scale_linetype_manual(values=c(1,1), name="Site Type")  + 
  scale_color_manual(values=c("#018571","#f8ab5b","#4fc1af","#a6611a"), name="Sites") +
  labs(x="PCoA Dim 1",y="PCoA Dim 2") +
  theme_bw() + theme(axis.text=element_blank(),axis.ticks=element_blank(),legend.text=element_text(size=14),
                     legend.title=element_text(size=14))

```

```{r echo=TRUE, eval=writing}
png(here('figs','Figure3_PCoA.png'),res=300,height=1800,width=3800)
plot_grid(fig3a,fig3b,rel_widths=c(0.72,1),labels=c("a","b"))
dev.off()
```

```{r echo=TRUE, eval=writing}
tiff(here('figs','Fig3.tif'),res=300,height=1800,width=3800)
plot_grid(fig3a,fig3b,rel_widths=c(0.72,1),labels=c("a","b"))
dev.off()
```
