---
title: "Sample Taxonomy Clean Up CEG Server BLAST"
author: "M Fisher"
subtitle: "Written: 2023-1-1"
date: 'Last run `R Sys.date()`'
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Description

Clean up the taxonomic information for each crab (+technical replicate). This involves:

1. Removing the ASVs in the positive and negative controls

2. Removing predator DNA

3. Removing non-target bacteria, fungal DNA

4. Removing DNA from species too small to be targeted prey, e.g., copepods, etc.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(here)
library(ggplot2)
library(ggrepel)

writing=TRUE  ## true to write files, false to knit for doc
```

Specify user directories
```{r}
blast_dir   <- 'data/blast'
outdir      <- 'data/blast'
```
<br>

What are the names of the processed blast file with taxonomy from script 6?
```{r}
# hashes with unique staxids
blast_file <- 'run4_bf3_sample_CEGblast_lca_taxonomy.csv'
```
<br>

What is the location / name of the metadata file for this sequencing run?
```{r}
run_metadat_file <- 'data/metadata/run4_BF3_filtered_metadata.csv'
```
<br>

And the run number and the marker
```{r}
run.num <- 4
marker <- 'BF3'
```
<br>

# Read in data

```{r read in asvs1, include=FALSE}
blast.dat <- read_csv(here(blast_dir, blast_file))
metadat <- read_csv(here(run_metadat_file))
```
<br>

Clean up the blast data, separate the sample-rep into a sample ID and replicate number, and then join to metadata information.
```{r}
blast.dat.filter <- blast.dat %>%
  dplyr::select(1:14) %>%
  mutate(type=ifelse(grepl("Positive", sample_id) | grepl("PCRNegative",sample_id) | grepl("ENegative",sample_id) | grepl("Negative",sample_id), "Control",
                     ifelse(grepl("MC",sample_id),"Mock Community","Sample"))) %>%
  separate(col=sample_id, into=c("sample","tech"), sep=-1,remove=FALSE) %>%
  mutate(tech=as.numeric(as.factor(tolower(tech)))) %>%
  mutate(sample=ifelse(type=="Sample" | type=="Mock Community",str_sub(sample,start=1,end=-2),sample),
         tech=ifelse(type=="Sample" | type=="Mock Community",tech,NA)) %>%
  mutate(sample_id=as.character(sample_id))

# blast.dat.filter <- left_join(blast.dat.filter, metadat,by=c("sample_id"="Sample_label","Sample_name","Locus"))
blast.dat.filter <- left_join(blast.dat.filter, metadat,by=c("sample_id"="Sample_label","Sample_name","sample","tech"))
dim(blast.dat.filter)
```
<br>
<br>

# No Match

What proportion of Hashes have no taxonomic information available?

```{r}
blast.dat.filter %>%
  dplyr::select(sample,Hash,rank) %>%
  distinct() %>%
  group_by(sample) %>%
  summarise(total_hashes=n(),
            n_missing=sum(is.na(rank))) %>%
  mutate(p_missing=n_missing/total_hashes) %>%
  ggplot(aes(x=sample,y=p_missing)) + geom_col() + 
  labs(y="Prop Hashes No Taxonomy") + ggtitle(paste0("Run ",run.num)) +
  ylim(c(0,1)) + 
  theme_bw() + theme(axis.text.x=element_blank())
```

```{r}
blast.dat.filter %<>% filter(!is.na(rank))
```
<br>


# Identify Taxonomy for Removal

To easier track how much data is being removed, I'm going to classify each hash as (1) to remove, and (2) reason for removal. then in the next section, I'll filter the data and check out the prey taxa. 

### Controls

get the hashes in the controls. **Not run through BLAST for Run 2, set to eval=FALSE**
```{r echo=FALSE, fig.height=3, fig.width=8, eval=FALSE}
control_dat <- blast.dat.filter %>% filter(type=="Control")
ggplot(data=control_dat, aes(x=sample_id, y=nReads, fill=taxon)) +
  geom_col() +
  facet_wrap(~sample_id, scales="free_y", ) +
  theme_bw() + ggtitle("Controls - Taxa") + theme(axis.text.x=element_blank(),axis.title.x=element_blank())
```
<br>
```{r echo=FALSE}
# control_dat <- blast.dat.filter %>% filter(type=="Control" | taxon %in% c("Homo sapiens","Primates"))

control_dat <- blast.dat.filter %>% filter((type=="Control" & taxon != "Carcinus maenas"))
```
<br>

mark in the rest of the data. 
```{r}
blast.dat.filter <- blast.dat.filter %>% 
  filter(type=="Sample" | type=="Mock Community") %>%
  mutate(control_rm=ifelse(Hash %in% control_dat$Hash,1,0))
```
<br>

there was also some human DNA (oops!) in some runs that needs to be removed
```{r}
blast.dat.filter %<>%
  mutate(human_contamination=ifelse(control_rm==0 & taxon %in% c("Homo sapiens","Primates"), 1, 0))
```

### Predator DNA

Any Green crab DNA will be marked as 'predator' DNA and removed. 

- Carcinus maenas (species)
- Carcinus (genus)
<br>

```{r}
blast.dat.filter <- blast.dat.filter %>% 
  mutate(predator_rm=ifelse(control_rm==0 & human_contamination==0 & (taxon =="Carcinus maenas" | taxon =="Carcinus aestuarii" | taxon =="Carcinus"), 1, 0))
```
<br>
<br>

### Bacterial, fungal etc. DNA

The data has the following phylums; and when phylum is missing, these are the classes :
```{r echo=FALSE}
unique(blast.dat.filter$phylum)
unique(filter(blast.dat.filter, is.na(phylum))$class)
```
<br>

Across all runs, there is also one genus of protists, and a genus of amoebas (not shown above).

These are the bacterial / fungal / amoeba / protist taxa to remove: 
```{r}
to_remove_phyla <- c("Proteobacteria","Planctomycetes","Planctomycetota","Oomycota","Actinobacteria","Basidiomycota","Ascomycota","Discosea","Evosea","Pseudomonadota","Actinomycetota")
to_remove_taxa <- c("Paraphysomonas sp.", "Amoebozoa sp.")

blast.dat.filter <- blast.dat.filter %>% 
  mutate(baf_rm=ifelse(control_rm==0 & human_contamination==0 & (phylum %in% to_remove_phyla | taxon %in% to_remove_taxa), 1, 0))
``` 
<br>
<br>

### Unlikely prey

Some items that left behind DNA in the crab stomach are unlikely to be targeted prey -- they could have been ingested while the crab was eating something else, or they could have been in the stomach of the crab's prey. 

I'm going to assume that the following are not targeted by the crab as prey: 

- diatoms (Bacillariophyta)
- rotifers (Rotifera)
- small algae & plankton (including golden algae)
- mackerel (genus: Scomber) <-- bait!

Here are the algae / plant taxa:
```{r echo=FALSE}
print(blast.dat.filter %>% filter(phylum %in% c("Rhodophyta","Chlorophyta","Streptophyta") | class %in% c("Phaeophyceae","Chrysophyceae")) %>% dplyr::select(phylum, class, order, taxon) %>% distinct())
```
<br>

From that list, I'm going to remove the following:
```{r}
to_remove_phyla <- c("Bacillariophyta", "Rotifera")
to_remove_taxa  <- c("uncultured Bangiophyceae","Tetraselmis","Chromulinales","Chloropicon","Chloropicon primus","Chloroparvula pacifica","	Paraphysomonas sp.","uncultured Jiaozhou Bay zooplankton","Trebouxia aggregata","Lobosphaera incisa",	"Scenedesmus quadricauda","Trebouxiophyceae sp. M-hakoo 311","Pseudoralfsia verrucosa","Thecamonas trahens","Cyprideis torosa","Gallus gallus","Myrionema strangulans","Apoikiospumella mondseeiensis","Pedospumella sp.","Synechococcus")
to_remove_genus <- c("Scomber","Chloropicon")
to_remove_class <- c("Chrysophyceae")

blast.dat.filter <- blast.dat.filter %>% 
  mutate(too_small_rm=ifelse(control_rm==0 & human_contamination==0 & (phylum %in% to_remove_phyla | taxon %in% to_remove_taxa | genus %in% to_remove_genus) | class %in% to_remove_class, 1, 0))
```
<br>

There are probably a range of other organisms that are not target prey, but I'm going to leave everything else in for now until I have a crab biologist weigh in. 
<br>

### Mock community
Are there any hashes from the mock communities in the other samples? Could this be tag jumping?
```{r}
mc_dat <- blast.dat.filter %>% filter(type=="Mock Community")
blast.dat.filter <- blast.dat.filter %>% 
  mutate(mc_rm=ifelse(Hash %in% mc_dat$Hash,1,0))
```


## Summary

Look at the proportion of reads in each 'removal' category, for each sample.
```{r echo=FALSE}
summary_dat <- blast.dat.filter %>%
  mutate(missing_rm=ifelse(is.na(rank) & is.na(taxon), 1,0))%>%
  mutate(retained=ifelse(missing_rm==0 & control_rm==0 & predator_rm==0 & baf_rm==0 & too_small_rm==0, 1,0)) %>%
  pivot_longer(cols=c(missing_rm, control_rm,predator_rm,baf_rm,too_small_rm,retained), names_to="reason") %>%
  mutate(reason=ifelse(reason=="missing_rm","missing",ifelse(reason=="control_rm","in control",
                                                             ifelse(reason=="predator_rm","predator DNA",
                                                                    ifelse(reason=="baf_rm","non-target DNA",
                                                                           ifelse(reason=="too_small_rm","incidental ingestion",ifelse(reason=="retained","retained",NA))))))) %>%
  filter(value==1) %>%
  dplyr::select(sample_id,sample,tech,site_month,Hash,nReads,reason,rank,taxon)

total_reads_per_sample <- blast.dat.filter %>%
  group_by(sample_id) %>%
  summarise(total_reads=sum(nReads), .groups="drop")

plotdat <- summary_dat %>%
  left_join(total_reads_per_sample,by="sample_id") %>%
  group_by(sample_id,site_month,sample,reason) %>%
  summarise(pReads=nReads/total_reads, .groups="drop")
```
<br>

Overall summary, by site: 
```{r fig.width=10,fig.height=6}
ggplot(plotdat %>% filter(site_month !="Mock Community"), aes(x=sample_id,y=pReads,fill=reason)) +
  geom_col() +
  facet_grid(cols=vars(site_month), scales="free") +
  xlab("crab + technical replicate") +
  theme_bw() + theme(axis.text.x=element_blank(), legend.title=element_blank())
```
<br>

By sampling site & month, with crab details:
```{r echo=FALSE, fig.height=5, fig.width=8}
site <- "Long Beach"
ggplot(plotdat %>% filter(grepl(site, site_month)), aes(x=sample_id,y=pReads,fill=reason)) +
  geom_col() +
  geom_vline(xintercept=seq(from=3.5, by=3, length.out=5)) +
  labs(x="crab + technical replicate", title=paste0(site)) +
  theme_bw() + theme(legend.title=element_blank(),
                     axis.text.x=element_text(angle=90,vjust=0.5,hjust=1))
```

```{r echo=FALSE, fig.height=5, fig.width=10}
site <- "Oysterville"
ggplot(plotdat %>% filter(grepl(site, site_month)), aes(x=sample_id,y=pReads,fill=reason)) +
  geom_col() +
  facet_grid(cols=vars(sample), scales="free") +
  labs(x="crab + technical replicate") + ggtitle(site) +
  theme_bw() + theme(legend.title=element_blank(),
                     axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),
                     strip.text=element_blank())
```
<br>

*No Nahcotta in runs 2,3 - set eval to FALSE*
```{r echo=FALSE, fig.height=5, fig.width=10, eval=FALSE}
site <- "Nahcotta"
ggplot(plotdat %>% filter(grepl(site, site_month)), aes(x=sample_id,y=pReads,fill=reason)) +
  geom_col() +
  facet_grid(cols=vars(sample), scales="free") +
  labs(x="crab + technical replicate") + ggtitle(site) +
  theme_bw() + theme(legend.title=element_blank(),
                     axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),
                     strip.text=element_blank())
```

What about the mock communities?
```{r echo=FALSE, fig.height=5, fig.width=8}
plotdat %>% filter(grepl("MC",sample_id)) %>%
ggplot(aes(x=sample_id,y=pReads,fill=reason)) +
  geom_col() +
  geom_vline(xintercept=seq(from=3.5, by=3, length.out=5)) +
  labs(x="community + technical replicate", title="Mock Community") +
  theme_bw() + theme(legend.title=element_blank(),
                     axis.text.x=element_text(angle=90,vjust=0.5,hjust=1))
```
<br>

# Write out Taxonomy

first, save a record of what / why ASVs were filtered
```{r}
blast.dat.filter %>%
  filter(!is.na(rank)) %>%
  # collapse non-target ASvs
  mutate(nonTarget_rm=ifelse(baf_rm==1 | too_small_rm==1, 1, 0)) %>%
  # get unique ASVs
  group_by(Hash,rank,kingdom,phylum,class,order,family,genus,species,taxon,human_contamination,control_rm,predator_rm,nonTarget_rm,type) %>%
  summarise(n.crab=length(unique(sample)),
          reads.in.run=sum(nReads)) %>%
  mutate(MiSeqRun=run.num) %>%
  write_csv(here(outdir,paste0("run",run.num,"_",marker,"_taxonomy_track_filter.csv")))
```




remove all of the hashes marked for removal above. 
```{r}
blast.dat.out <- blast.dat.filter %>%
  filter(!is.na(rank)) %>%
  filter(control_rm==0 & human_contamination==0 & predator_rm==0 & baf_rm==0 & too_small_rm==0) %>%
  dplyr::select(-c(control_rm,human_contamination,predator_rm,baf_rm,too_small_rm))
```
<br>

write out. 
```{r eval=writing, include=writing}
write_csv(blast.dat.out, here(outdir, paste0("run",run.num,"_",marker,"_taxonomy_filtered.csv")))
```
<br>

# Run Quality for Prey Info

How many samples have prey information? Of those, how many technical replicates were retained?
```{r}
prey_dat_summary <- blast.dat.out %>%
  group_by(sample,site_month) %>%
  summarise(n_reps=length(unique(tech)),
            n_prey_taxa=length(unique(taxon)),
            nReads_prey_taxa=sum(nReads), .groups="drop")

length(unique(metadat$sample))
length(unique(prey_dat_summary$sample))
```
<br>


Same as above, but without the Hashes that were in the mock community samples too. 
```{r eval=TRUE}
prey_dat_summary_noMC <- blast.dat.out %>%
  filter(mc_rm==0) %>%
  group_by(sample,site_month) %>%
  summarise(n_reps=length(unique(tech)),
            n_prey_taxa=length(unique(taxon)),
            nReads_prey_taxa=sum(nReads), .groups="drop")
```
<br>

Only specific prey IDs (genus or species). 
```{r}
# prey_specdat_IDs <- blast.dat.out %>%
#   filter(rank %in% c("species","genus")) %>%
#   dplyr::select(sample,site_month,rank,taxon,phylum,class,mc_rm) %>%
#   distinct()

prey_specdat_IDs <- blast.dat.out %>%
  filter(rank %in% c("species","genus")) %>%
  dplyr::select(sample,site_month,rank,taxon,phylum,class) %>%
  distinct()

prey_specdat_summary <- blast.dat.out %>% 
  filter(rank %in% c("species","genus")) %>%
  group_by(sample,site_month) %>%
  summarise(n_reps=length(unique(tech)),
            n_prey_taxa=length(unique(taxon)),
            nReads_prey_taxa=sum(nReads), .groups="drop") %>%
  left_join(prey_specdat_IDs) %>% distinct()
```
<br>

Save this.
```{r eval=writing, include=writing}
# prey_dat_summary_out <- prey_dat_summary %>%
#   mutate(filter_mcHash=0) %>%
#   bind_rows(prey_dat_summary_noMC %>%
#               mutate(filter_mcHash=1))

prey_dat_summary_out <- prey_dat_summary

write_csv(prey_dat_summary_out, here('data','qc', paste0("run",run.num,"_",marker,"_samples_with_prey_data.csv")))


write_csv(prey_specdat_summary, here(outdir, paste0("run",run.num,"_",marker,"_sample_prey_data_summary.csv")))
```







