---
title: "Pamphlet Graphics"
author: "M Fisher"
date: "2023-03-26"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Description

Create the graphics for the stakeholder pamphlet deliverable. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(janitor)
library(magrittr)
library(vegan)
library(ggmap)

source(here('R','get_unique_taxa.R'))

# User inputs
indir   <- 'results'
run.nums <- c(2,3,4,5)
marker  <- 'BF3'
metadata_dir <- 'data/metadata'
```
```{r data, include=FALSE}

dat <- read_csv(here('data','results','allRuns_BF3_filtered_FINAL_unique_taxa.csv'))

groups.key <- read_csv(here('data','results','taxa_groups_key.csv'))

# metadata filtered (crab with prey info)
for(r in run.nums){
  if(r==run.nums[1]){
    meta <- read.csv(here(metadata_dir,paste0("run",r,"_",marker,"_filtered_metadata.csv"))) %>% clean_names()}
  else{
    meta %<>% bind_rows(read.csv(here(metadata_dir,paste0("run",r,"_",marker,"_filtered_metadata.csv")))
                        %>% clean_names())
    
  }
}

site_coords <- read_csv(here('data','metadata','site_coords.csv'))
```
```{r eval=FALSE}
# some extra QC: any hashes that are listed more than once in nonspec for the same crab?
dat %>%
  filter(rank!="species") %>%
  dplyr::select(sample,Hash,taxon,rank,species,genus,family,order,class) %>% distinct() %>%
  group_by(Hash, sample) %>% summarise(duplicated.ranks=length(unique(rank))) %>% filter(duplicated.ranks > 1)
```
<br>

# Sample counts, metadata

```{r}
length(unique(dat$sample))
```


# Basic Prey Diversity

What was the unique number of taxa identified?

### Across all sites

Species:
```{r}
# get species-level prey for each crab
spdat <- dat %>% filter(rank=="species") %>% dplyr::select(taxon,genus,family,order,class) %>% distinct() %>%
  rename("sp.taxon"=taxon)

length(spdat$sp.taxon)
```

Unique taxa:
```{r}
tmp <- get_unique_taxa(dat) %>%
  dplyr::select(taxon,rank, phylum,class,order,family,genus,species) %>% distinct()

length(tmp$taxon)
```


### By site

Species
```{r}
dat %>% filter(rank=="species") %>% 
  separate(site_month,into=c("site","month"),sep="-",remove=TRUE) %>%
  dplyr::select(site,taxon,genus,family,order,class) %>% distinct() %>%
  group_by(site) %>% summarise(n.species=length(unique(taxon)))
```

Taxa
```{r}
dat_sites <- dat %>% separate(site_month, into=c("site","month"),sep="-") %>%
get_unique_taxa(level="site")

dat_sites %>%
  dplyr::select(taxon,site,rank, phylum,class,order,family,genus,species) %>%
  group_by(site) %>% summarise(n.taxa=length(unique(taxon)))
```


### By month
Species
```{r}
dat %>% filter(rank=="species") %>% 
  separate(site_month,into=c("site","month"),sep="-",remove=TRUE) %>%
  rename(site=month, x=site) %>%
  dplyr::select(site,taxon,genus,family,order,class) %>% distinct() %>%
  group_by(site) %>% summarise(n.species=length(unique(taxon))) %>%
  rename(month=site)
```

Taxa
```{r}
dat %>% separate(site_month,into=c("site","month"),sep="-",remove=TRUE) %>%
  rename(site=month, x=site) %>%
get_unique_taxa(level="site") %>%
  dplyr::select(taxon,site,rank, phylum,class,order,family,genus,species) %>%
  group_by(site) %>% summarise(n.taxa=length(unique(taxon))) %>%
  rename(month=site)
```

# Prey Identification

```{r}
total.crab <- length(unique(dat$sample))
crab.by.site.type <- dat %>% 
  separate(site_month,into=c("site","month"),sep="-",remove=TRUE) %>%
  mutate(site_type=ifelse(grepl("Nahcotta",site) | grepl("Stackpole",site),"Monitoring Sites","Clam Bed Sites")) %>%
  group_by(site_type) %>% summarise(site.n=length(unique(sample)))
```


### by group

```{r}
preyGrp.site <- dat %>%
  separate(site_month,into=c("site","month"),sep="-",remove=TRUE) %>%
  left_join(groups.key, by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  ### for the pamphlet, don't break out algae by group (red/green/brown) ###
  mutate(taxon.group=ifelse(grepl("algae",taxon.group), "Algae",taxon.group)) %>%
  group_by(site) %>%
  mutate(site_reads=sum(nReads)) %>%
  ungroup() %>%
  group_by(site, taxon.group) %>% 
  summarise(n_crab=length(unique(sample)),
            n_reads=sum(nReads),
            prop_reads=n_reads/site_reads) %>% distinct() %>% arrange(site,n_crab) %>%
  mutate(site_type=ifelse(grepl("Nahcotta",site) | grepl("Stackpole",site),"Monitoring Sites","Clam Bed Sites")) %>%
  arrange(n_crab)

preyGrp.site$site <- factor(preyGrp.site$site, levels=c("Long Beach","Oysterville","Stackpole","Nahcotta"))
preyGrp.site$taxon.group <- factor(preyGrp.site$taxon.group, levels=rev(c("Algae","Other Crabs",
                                                                          "Shrimp","Amphipods","Fishes",
                                                                          "Worms","Insects","Bivalves",
                                                                          "Plants","Invasive mud snail",
                                                                          "Tunicates","Sea anemone")))
```
<br>
```{r}
col_palette3 <- c("#01665e","#5ab4ac","#c7eae5",
                  "#8c510a","#d8b365","#f6e8c3",
                  "#08519c","#9ecae1",
                  "#c51b7d","#e9a3c9")
col_palette2 <- c("#01665e","#c7eae5",
                  "#8c510a","#f6e8c3")


ggplot(preyGrp.site, aes(y=taxon.group,x=n_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + ggtitle("What are green crabs eating?") + labs(x="Number of green crabs") +
  scale_fill_manual(values=col_palette2) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12))
```

```{r}
png(here('figs','POSTER_prey-groups_counts-by-site.png'),res=300,height=1200,width=1600,bg="transparent")
ggplot(preyGrp.site, aes(y=taxon.group,x=n_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + labs(x="Number of green crabs") +
  scale_fill_manual(values=col_palette2) + guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12),
                     strip.text=element_text(size=13), legend.position="bottom",
                     panel.grid.major=element_line(linewidth=0.25),
                     # panel.background = element_rect(fill='transparent'), #transparent panel bg
                     plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                     legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                     legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```

```{r}
preyGrp.site %>%
  left_join(crab.by.site.type) %>%
  mutate(p_crab=100*(n_crab/site.n)) %>%
  ggplot(aes(y=taxon.group,x=p_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + ggtitle("What are green crabs eating?") + labs(x="Percent of crabs") +
  scale_fill_manual(values=col_palette2) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12))
```
```{r}
png(here('figs','POSTER_prey-groups_pcrabs-by-site.png'),res=300,height=1200,width=1600,bg="transparent")
preyGrp.site %>%
  left_join(crab.by.site.type) %>%
  mutate(p_crab=100*(n_crab/site.n)) %>%
  ggplot(aes(y=taxon.group,x=p_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + labs(x="Percent of green crabs") +
  scale_fill_manual(values=col_palette2) + guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12),
                     strip.text=element_text(size=13), legend.position="bottom",
                     panel.grid.major=element_line(linewidth=0.25),
                     # panel.background = element_rect(fill='transparent'), #transparent panel bg
                     plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                     legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                     legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```

### % crab per species

"top" species
```{r}
dat %>% left_join(groups.key,by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  group_by(rank,taxon,taxon.common) %>%
  summarise(n.crab=length(unique(sample))) %>%
  mutate(p.crab=n.crab/total.crab) %>%
  filter(!is.na(p.crab)) %>%
  arrange(desc(p.crab)) 
```

Combine the Crangon (species/genus) IDs and recalculate percent
```{r}
dat %>% left_join(groups.key,by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  filter(taxon %in% c("Crangon","Crangon franciscorum")) %>%
  mutate(taxon.group="Crangon shrimp") %>%
  group_by(taxon.group) %>%
  summarise(n.crab=length(unique(sample))) %>%
  mutate(p.crab=n.crab/total.crab)
```

Graph percentages for bivalves, crab, and fish (POSTER version)
```{r}
dat_sp <- dat %>% left_join(groups.key,by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  filter(!is.na(taxon.common)) %>%
  separate(site_month, into=c("site","month"), sep="-") %>%
  group_by(site,rank,taxon,taxon.common,phylum,class,order,family,taxon.group) %>%
  summarise(n.crab=length(unique(sample))) %>%
  left_join(dat %>% separate(site_month, into=c("site","month"), sep="-") %>%
              group_by(site) %>% 
              summarise(site.crab=length(unique(sample))), by="site") %>%
  mutate(p.crab=n.crab/site.crab*100) %>%
  arrange(desc(p.crab))


plotdat <- dat_sp %>% 
  rename(`Percent of Crabs\nwith food item`=p.crab) %>%
  mutate(site_type=ifelse(site %in% c("Long Beach","Oysterville"), "Clam Bed", "Monitoring")) %>%
  mutate(taxon.common=ifelse(taxon.common=="Shiner perch (C.aggregata)","Shiner perch (C. aggregata)",taxon.common))



plotdat$taxon.common <- factor(plotdat$taxon.common, levels=rev(c("Soft-shell clam (M. arenaria)","Manila clam",
                                                                  "Shore crab (H. oregonensis)", "Shore crab (Hemigrapsus spp.)","Dungeness crab",
                                                                  "Sand shrimp (C. franciscorum)", "Shrimp (Crangon spp.)",
                                                                  "Shiner perch (C. aggregata)","Pacific staghorn sculpin (L. armatus)","Saddleback gunnel (P. ornata)")))

plotdat$site <- factor(plotdat$site, levels=c("Long Beach","Oysterville","Stackpole","Nahcotta"))

col_palette4 <- c("#018571","#a6611a")
col_palette2 <- c("#01665e","#c7eae5",
                  "#8c510a","#f6e8c3")
plotdat %>% filter(!is.na(taxon.common)) %>%
ggplot() +
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`, col=site)) + 
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`), pch=21) +
  geom_rect(aes(xmin=9.5,xmax=8.5,ymin=1.75,ymax=2.25), col="gray60",fill=NA) +
  geom_segment(aes(x=9,xend=8.80,y=2.25,yend=2.4), col="gray60") +
  geom_text(data=data.frame(x=8.9,y=3.12,label="1 green crab"), aes(x=x,y=y,label=label), col="gray60", fontface="italic",size=4) +
  scale_color_manual(values=col_palette2,name="") +
  scale_size_continuous(range=c(2,6)) +
  guides(color=FALSE) +
  labs(x="",y="") + ggtitle("What bivalves, fish, and other crab are being eaten?") +
  coord_flip() + theme_bw() + theme(axis.text.x=element_text(angle=20,hjust=1,size=12),
                                    axis.text.y=element_text(size=12),
                                    legend.text=element_text(size=12), legend.title=element_text(size=13))
```
```{r}
png(here('figs','POSTER_prey-clams-crab-fish_by-site_dotplot.png'),res=300,height=1500,width=2000,bg = 'transparent')
plotdat %>% filter(!is.na(taxon.common)) %>%
ggplot() +
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`, col=site)) + 
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`), pch=21) +
  geom_rect(aes(xmin=9.5,xmax=8.5,ymin=1.75,ymax=2.25), col="gray60",fill=NA) +
  geom_segment(aes(x=9,xend=8.80,y=2.25,yend=2.4), col="gray60") +
  geom_text(data=data.frame(x=8.9,y=3.32,label="1 green crab"), aes(x=x,y=y,label=label), col="gray60", fontface="italic",size=4) +
  scale_color_manual(values=col_palette2,name="") +
  scale_size_continuous(range=c(2,6)) +
  guides(color=FALSE) +
  labs(x="",y="") + ggtitle("What bivalves, fishes, and other\n crabs are being eaten?") +
  coord_flip() + theme_bw() + theme(axis.text.x=element_text(angle=30,hjust=1,size=12),
                                    axis.text.y=element_text(size=12),
                                    legend.text=element_text(size=12), legend.title=element_text(size=13),
                                    panel.grid.major=element_line(linewidth=0.25),
                                    # panel.background = element_rect(fill='transparent'), #transparent panel bg
                                    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                                    legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                                    legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```



Graph percentages for bivalves, crab, and fish (PAMPHLET version)
```{r}
dat_sp2 <- dat %>% left_join(groups.key,by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  # add kurtiella clam common name
  mutate(taxon.common=ifelse(grepl("Kurtiella",taxon), "Kurtiella sp.\n(small clam)", taxon.common)) %>%
  # only keep taxa with common names
  filter(!is.na(taxon.common)) %>%
  filter(!grepl("Crangon",taxon) & taxon!="Hemigrapsus") %>%   # get rid of genus-level IDs of these two groups
  # remove scientific names
  mutate(taxon.common=ifelse(taxon.common=="Soft-shell clam (M. arenaria)", "Soft-shell clam",
                             ifelse(taxon.common=="Shiner perch (C. aggregata)","Shiner perch",
                                    ifelse(taxon.common=="Pacific staghorn sculpin (L. armatus)","Pacific staghorn\n sculpin",
                                           ifelse(taxon.common=="Saddleback gunnel (P. ornata)","Saddleback gunnel",
                                                  ifelse(taxon.common=="Shore crab (H. oregonensis)","Hairy shore crab",
                                           taxon.common)))))) %>%
  separate(site_month, into=c("site","month"), sep="-") %>%
  group_by(site,rank,taxon,taxon.common,phylum,class,order,family,taxon.group) %>%
  summarise(n.crab=length(unique(sample))) %>%
  left_join(dat %>% separate(site_month, into=c("site","month"), sep="-") %>%
              group_by(site) %>% 
              summarise(site.crab=length(unique(sample))), by="site") %>%
  mutate(p.crab=n.crab/site.crab*100) %>%
  arrange(desc(p.crab))


plotdat <- dat_sp2 %>% 
  rename(`Percent of\ngreen crabs\nwith food item`=p.crab) %>%   ## rename legend
  mutate(site_type=ifelse(site %in% c("Long Beach","Oysterville"), "Clam Bed", "Monitoring")) %>%
  mutate(taxon.common=ifelse(taxon.common=="Shiner perch (C.aggregata)","Shiner perch (C. aggregata)",taxon.common))



plotdat$taxon.common <- factor(plotdat$taxon.common, levels=rev(c("Soft-shell clam","Manila clam","Kurtiella sp.\n(small clam)",
                                                                  "Shiner perch","Pacific staghorn\n sculpin","Saddleback gunnel",
                                                                  "Hairy shore crab", "Dungeness crab")))

plotdat$site <- factor(plotdat$site, levels=c("Long Beach","Oysterville","Stackpole","Nahcotta"))

col_palette4 <- c("#018571","#a6611a")
col_palette2 <- c("#01665e","#c7eae5",
                  "#8c510a","#f6e8c3")
plotdat %>% filter(!is.na(taxon.common)) %>%
ggplot() +
  geom_point(aes(x=taxon.common,y=site,size=`Percent of\ngreen crabs\nwith food item`, col=site)) + 
  geom_point(aes(x=taxon.common,y=site,size=`Percent of\ngreen crabs\nwith food item`), pch=21) +
  geom_rect(aes(xmin=7.5,xmax=6.5,ymin=1.75,ymax=2.25), col="gray60",fill=NA) +
  geom_segment(aes(x=7,xend=6.80,y=2.25,yend=2.4), col="gray60") +
  geom_text(data=data.frame(x=6.4,y=3.12,label="1 green crab"), aes(x=x,y=y,label=label), col="gray60", fontface="italic",size=4) +
  scale_color_manual(values=col_palette2,name="") +
  scale_size_continuous(range=c(2,6)) +
  guides(color=FALSE) +
  labs(x="",y="") + ggtitle("What bivalves, fish, and other crabs are being eaten?") +
  coord_flip() + theme_bw() + theme(axis.text.x=element_text(angle=20,hjust=1,size=12),
                                    axis.text.y=element_text(size=12),
                                    legend.text=element_text(size=12), legend.title=element_text(size=13))
```
```{r}
png(here('figs','PAMPHLET_prey-clams-crab-fish_by-site_dotplot.png'),res=300,height=1500,width=1600,bg = 'transparent')
plotdat %>% filter(!is.na(taxon.common)) %>%
  ggplot() +
  geom_point(aes(x=taxon.common,y=site,size=`Percent of\ngreen crabs\nwith food item`, col=site)) + 
  geom_point(aes(x=taxon.common,y=site,size=`Percent of\ngreen crabs\nwith food item`), pch=21) +
  geom_rect(aes(xmin=7.5,xmax=6.5,ymin=1.75,ymax=2.25), col="gray60",fill=NA) +
  geom_segment(aes(x=7,xend=6.80,y=2.25,yend=2.4), col="gray60") +
  geom_text(data=data.frame(x=6.82,y=3.25,label="1 green crab"), aes(x=x,y=y,label=label), col="gray60", fontface="italic",size=4) +
  scale_color_manual(values=col_palette2,name="") +
  scale_size_continuous(range=c(2,6)) +
  guides(color=FALSE) +
  labs(x="",y="") + ggtitle("What bivalves, fishes, and other\n crabs are being eaten?") +
  coord_flip() + theme_bw()+ theme(axis.text.x=element_text(angle=30,hjust=1,size=12),
                                   axis.text.y=element_text(size=12),
                                   legend.text=element_text(size=12), legend.title=element_text(size=13),
                                   panel.grid.major=element_line(linewidth=0.25),
                                   # panel.background = element_rect(fill='transparent'), #transparent panel bg
                                   plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                                   legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                                   legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```



# Map

### Site Map

data needed for mapping
```{r}
willapa <- readRDS(here("data","metadata","willapa_terrain_background.rds"))
mapdat <- site_coords %>% dplyr::select(site_name,latitude,longitude,type) %>% distinct() %>%
  mutate(type=ifelse(type=="Sentinel","Monitoring\nSites","Clam Bed\nSites"))
```

create map
```{r}
mymap2 <- ggmap(willapa) + 
  geom_label_repel(data=mapdat, aes(x=longitude,y=latitude, label=site_name, color=type),
                   segment.colour = 'black',
                   nudge_x=0.2,
                   force=1, force_pull=0.2, segment.size=0.05, size=5) +
  geom_point(data=mapdat, aes(x=longitude,y=latitude, col=type), size=6) +
  geom_text(aes(x=-123.98,y=46.65,label="italic('Willapa\n Bay')"), col="deepskyblue4", size=4, parse=TRUE) +
  geom_text(aes(x=-123.85,y=46.23,label="italic('Columbia River')"), col="deepskyblue4", size=4, parse=TRUE) +
  scale_color_manual(values=c("#018571","#a6611a")) +
  theme(axis.title=element_blank(),
        axis.text=element_blank(), axis.ticks=element_blank(),
        legend.title=element_blank(),
        legend.text=element_text(size=12), legend.position="right",legend.justification='top',
        plot.margin=unit(c(0.5,0,0,0), "cm"))
mymap2
```



*Alternative*
```{r}
willapa2 <- readRDS(here("data","metadata","willapa_terrain_background2.rds"))
mymap <- ggmap(willapa2) +
  geom_label_repel(data=mapdat, aes(x=longitude,y=latitude, label=site_name, color=type),
                   segment.colour = 'black',
                   nudge_x=0.2,
                   force=1, force_pull=0.2, segment.size=0.05, size=4) +
  geom_point(data=mapdat, aes(x=longitude,y=latitude, col=type), size=6) +
  geom_text(aes(x=-123.98,y=46.65,label="italic('Willapa\n Bay')"), col="deepskyblue4", size=4, parse=TRUE) +
  geom_text(aes(x=-123.87,y=46.23,label="italic('Columbia River')"), col="deepskyblue4", size=4, parse=TRUE) +
  ylab("Latitude") + xlab("Longitude") +
  scale_color_manual(values=c("#018571","#a6611a")) +
  theme(axis.title=element_blank(),
        axis.text=element_text(size=10),
        legend.title=element_blank(),
        legend.text=element_text(size=12), legend.position="right",legend.justification='top',
        plot.margin=unit(c(0,0,0,0), "cm")) +
  coord_fixed(ylim=c(46.3,46.8), xlim=c(-124.2, -123.7))
mymap
```


### Carapace widths (mm)

**Note that "sequenced" here is "sequenced with prey", whereas in the publication, there is a distinction between "sequenced" (n=86, inclusive of crabs which did not have identifiable prey in their stomach contents) and "sequenced with prey"**

```{r}
cw <- read_csv(here('data','metadata','Willapa Bay EGC Samples - Sample Data.csv')) %>%
  mutate(sequenced=ifelse(Sample_label %in% dat$sample, "Sequenced", "Trapped"))

cw %<>% mutate(site=ifelse(grepl("Stackpole",Site_Name), "Stackpole",
                                 ifelse(grepl("Oysterville",Site_Name), "Oysterville",
                                        ifelse(grepl("Nahcotta",Site_Name),"Nahcotta",
                                               ifelse(grepl("Long Beach",Site_Name),"Long Beach", NA))))) %>%
  mutate(site_type=ifelse(site %in% c("Stackpole","Nahcotta"),"Monitoring Sites","Clam Bed Sites"))

cw %>% group_by(sequenced, Sex) %>% summarise(ncrab=length(unique(Sample_label)))
```


```{r fig.width=6, fig.height=3}
cw.sequenced <- cw%>%filter(sequenced=="Sequenced")
plotdat <- cw %>% mutate(sequenced="Trapped") %>%
  bind_rows(cw.sequenced)

plotdat$site_type  <- factor(plotdat$site_type, levels=c("Clam Bed Sites","Monitoring Sites"))

ggplot(plotdat, aes(x=CW_mm, fill=sequenced)) +
  geom_histogram(position="dodge") +
  scale_fill_manual(values=c("black","grey60"),name="") +
  facet_grid(rows=vars(site_type),switch="y") + 
  scale_y_continuous(position = "right") +
  labs(x="Carapace Width (mm)",y="Number of Green Crabs") + theme_bw() + 
  theme(legend.position="top",strip.placement = "outside",
        strip.background=element_rect(color=c("white"), linewidth=2))
```

Alternative legend location
```{r}

ggplot(plotdat, aes(x=CW_mm, fill=sequenced)) +
  geom_histogram(position="dodge") +
  scale_fill_manual(values=c("black","grey60"),name="") +
  facet_grid(rows=vars(site_type),switch="y") + 
  scale_y_continuous(position = "right") +
  labs(x="Carapace Width (mm)",y="Number of green crabs") + theme_bw() + 
  theme(strip.placement = "outside",legend.position=c(0.1,0.9), legend.title=element_blank(),
        strip.background=element_rect(color=c("white"), linewidth=2))

plot1 <- ggplot(plotdat, aes(x=CW_mm, fill=sequenced)) +
  geom_histogram(position="dodge") +
  scale_fill_manual(values=c("black","grey60"),name="") +
  facet_grid(rows=vars(site_type),switch="y") + 
  scale_y_continuous(position = "right") +
  labs(x="Carapace width (mm)",y="Number of green crabs") + theme_bw() + 
  theme(strip.placement = "outside",legend.position=c(0.4,0.9), legend.title=element_blank(),
        strip.text=element_text(size=12),
        strip.background=element_rect(color=c("white"), linewidth=2))
```



### combined

```{r echo=TRUE}
right_side <- plot_grid(NULL, plot1, NULL, nrow=3, rel_heights=c(0.2,1,0.2))
png(here('figs','PAMPHLET_map-metadata.png'),res=300, height=2300,width=2300)
plot_grid(mymap2,right_side,ncol=2, rel_widths=c(1,0.5))
dev.off()


# png(here('figs','PAMPHLET_map-metadata-largerCW.png'),res=300, height=2000,width=2200)
# plot_grid(mymap2,plot1,ncol=2, rel_widths=c(1,0.7))
# dev.off()
```
