---
title: "Pamphlet Graphics"
author: "M Fisher"
date: "2023-03-26"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Description

Create the graphics for the stakeholder pamphlet deliverable. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
library(tidyverse)
library(here)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(janitor)
library(magrittr)
library(vegan)
library(ggmap)

source(here('R','get_unique_taxa.R'))

# User inputs
indir   <- 'results'
run.nums <- c(2,3,4,5)
marker  <- 'BF3'
metadata_dir <- 'data/metadata'
```
```{r data, include=FALSE}

dat <- read_csv(here('data','results','allRuns_BF3_filtered_FINAL_unique_taxa.csv'))

groups.key <- read_csv(here('data','results','taxa_groups_key.csv'))

# metadata filtered (crab with prey info)
for(r in run.nums){
  if(r==run.nums[1]){
    meta <- read.csv(here(metadata_dir,paste0("run",r,"_",marker,"_filtered_metadata.csv"))) %>% clean_names()}
  else{
    meta %<>% bind_rows(read.csv(here(metadata_dir,paste0("run",r,"_",marker,"_filtered_metadata.csv")))
                        %>% clean_names())
    
  }
}

site_coords <- read_csv(here('data','metadata','site_coords.csv'))
```
```{r eval=FALSE}
# some extra QC: any hashes that are listed more than once in nonspec for the same crab?
dat %>%
  filter(rank!="species") %>%
  dplyr::select(sample,Hash,taxon,rank,species,genus,family,order,class) %>% distinct() %>%
  group_by(Hash, sample) %>% summarise(duplicated.ranks=length(unique(rank))) %>% filter(duplicated.ranks > 1)
```
<br>


# Basic Prey Diversity

What was the unique number of taxa identified?

### Across all sites

Species:
```{r}
# get species-level prey for each crab
spdat <- dat %>% filter(rank=="species") %>% dplyr::select(taxon,genus,family,order,class) %>% distinct() %>%
  rename("sp.taxon"=taxon)

length(spdat$sp.taxon)
```

Unique taxa:
```{r}
tmp <- get_unique_taxa(dat) %>%
  dplyr::select(taxon,rank, phylum,class,order,family,genus,species) %>% distinct()

length(tmp$taxon)
```


### By site

Species
```{r}
dat %>% filter(rank=="species") %>% 
  separate(site_month,into=c("site","month"),sep="-",remove=TRUE) %>%
  dplyr::select(site,taxon,genus,family,order,class) %>% distinct() %>%
  group_by(site) %>% summarise(n.species=length(unique(taxon)))
```

Taxa
```{r}
dat_sites <- dat %>% separate(site_month, into=c("site","month"),sep="-") %>%
get_unique_taxa(level="site")

dat_sites %>%
  dplyr::select(taxon,site,rank, phylum,class,order,family,genus,species) %>%
  group_by(site) %>% summarise(n.species=length(unique(taxon)))


```


### By month
```{r}

```

# Beta diversity
What is the $\beta$ diversity for prey taxon? (the ratio between regional and local prey diversity) The greater the similarity in community composition between multiple communities, the lower the value of β-diversity for that set of communities.
```{r}
prey.sp.site.mat <- dat_sites %>%
  dplyr::select(site, taxon) %>% distinct() %>%
  mutate(presence=1) %>%
  pivot_wider(id_cols="site", names_from="taxon",values_from="presence", values_fill=0)

bdiv <- as.matrix(betadiver(prey.sp.site.mat, method="w")) 
colnames(bdiv) <-c(prey.sp.site.mat$site); rownames(bdiv) <- c(prey.sp.site.mat$site)
bdiv
```


# Prey Identification

```{r}
total.crab <- length(unique(dat$sample))
crab.by.site.type <- dat_sites %>% mutate(site_type=ifelse(grepl("Nahcotta",site) | grepl("Stackpole",site),"Monitoring Sites","Clam Bed Sites")) %>%
  group_by(site_type) %>% summarise(site.n=length(unique(sample)))
```


### by group

```{r}
preyGrp.site <- dat_sites %>%
  left_join(groups.key, by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  group_by(site) %>%
  mutate(site_reads=sum(nReads)) %>%
  ungroup() %>%
  group_by(site, taxon.group) %>% 
  summarise(n_crab=length(unique(sample)),
            n_reads=sum(nReads),
            prop_reads=n_reads/site_reads) %>% distinct() %>% arrange(site,n_crab) %>%
  mutate(site_type=ifelse(grepl("Nahcotta",site) | grepl("Stackpole",site),"Monitoring Sites","Clam Bed Sites")) %>%
  arrange(n_crab)

preyGrp.site$site <- factor(preyGrp.site$site, levels=c("Long Beach","Oysterville","Stackpole","Nahcotta"))
preyGrp.site$taxon.group <- factor(preyGrp.site$taxon.group, levels=rev(c("Other Crabs","Shrimp","Amphipods",
                                                                          "Bivalves",
                                                                          "Brown algae","Green algae","Red algae",
                                                                          "Worms","Insects","Plants",
                                                                          "Fishes","Tunicates","Invasive mud snail","Sea anemone")))
```
<br>
```{r}
col_palette3 <- c("#01665e","#5ab4ac","#c7eae5",
                  "#8c510a","#d8b365","#f6e8c3",
                  "#08519c","#9ecae1",
                  "#c51b7d","#e9a3c9")
col_palette2 <- c("#01665e","#c7eae5",
                  "#8c510a","#f6e8c3")


ggplot(preyGrp.site, aes(y=taxon.group,x=n_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + ggtitle("What are green crabs eating?") + labs(x="Number of crabs") +
  scale_fill_manual(values=col_palette2) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12))
```

```{r}
png(here('figs','POSTER_prey-groups_counts-by-site.png'),res=300,height=1200,width=1600,bg="transparent")
ggplot(preyGrp.site, aes(y=taxon.group,x=n_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + labs(x="Number of crabs") +
  scale_fill_manual(values=col_palette2) + guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12),
                     strip.text=element_text(size=13), legend.position="bottom",
                     panel.grid.major=element_line(linewidth=0.25),
                     # panel.background = element_rect(fill='transparent'), #transparent panel bg
                     plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                     legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                     legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```

```{r}
preyGrp.site %>%
  left_join(crab.by.site.type) %>%
  mutate(p_crab=100*(n_crab/site.n)) %>%
  ggplot(aes(y=taxon.group,x=p_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + ggtitle("What are green crabs eating?") + labs(x="Percent of crabs") +
  scale_fill_manual(values=col_palette2) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12))
```
```{r}
png(here('figs','POSTER_prey-groups_pcrabs-by-site.png'),res=300,height=1200,width=1600,bg="transparent")
preyGrp.site %>%
  left_join(crab.by.site.type) %>%
  mutate(p_crab=100*(n_crab/site.n)) %>%
  ggplot(aes(y=taxon.group,x=p_crab, fill=site)) + 
  geom_col() + facet_grid(cols=vars(site_type)) + labs(x="Percent of crabs") +
  scale_fill_manual(values=col_palette2) + guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme_bw() + theme(legend.title=element_blank(), axis.title.y=element_blank(),
                     axis.text.y=element_text(size=12), legend.text=element_text(size=12),
                     strip.text=element_text(size=13), legend.position="bottom",
                     panel.grid.major=element_line(linewidth=0.25),
                     # panel.background = element_rect(fill='transparent'), #transparent panel bg
                     plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                     legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                     legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```

### % crab per species
```{r}
dat_sites %>% left_join(groups.key,by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  group_by(rank,taxon,taxon.common) %>%
  summarise(n.crab=length(unique(sample))) %>%
  mutate(p.crab=n.crab/total.crab) %>%
  arrange(desc(p.crab)) %>% top_n(5)
```


For bivalves, crab, and fish



```{r}
dat_sp <- dat %>% left_join(groups.key,by=c("taxon"="taxon.final","rank"="rank.final")) %>%
  filter(!is.na(taxon.common)) %>%
  separate(site_month, into=c("site","month"), sep="-") %>%
  group_by(site,rank,taxon,taxon.common,phylum,class,order,family,taxon.group) %>%
  summarise(n.crab=length(unique(sample))) %>%
  left_join(dat %>% separate(site_month, into=c("site","month"), sep="-") %>%
              group_by(site) %>% 
              summarise(site.crab=length(unique(sample))), by="site") %>%
  mutate(p.crab=n.crab/site.crab*100) %>%
  arrange(desc(p.crab))


plotdat <- dat_sp %>% 
  rename(`Percent of Crabs\nwith food item`=p.crab) %>%
  mutate(site_type=ifelse(site %in% c("Long Beach","Oysterville"), "Clam Bed", "Monitoring")) %>%
  mutate(taxon.common=ifelse(taxon.common=="Shiner perch (C.aggregata)","Shiner perch (C. aggregata)",taxon.common))



plotdat$taxon.common <- factor(plotdat$taxon.common, levels=rev(c("Soft-shell clam (M. arenaria)","Manila clam",
                                                                  "Shore crab (H. oregonensis)", "Shore crab (Hemigrapsus spp.)","Dungeness crab",
                                                                  "Sand shrimp (C. franciscorum)", "Shrimp (Crangon spp.)",
                                                                  "Shiner perch (C. aggregata)","Pacific staghorn sculpin (L. armatus)","Saddleback gunnel (P. ornata)")))

plotdat$site <- factor(plotdat$site, levels=c("Long Beach","Oysterville","Stackpole","Nahcotta"))

col_palette4 <- c("#018571","#a6611a")
col_palette2 <- c("#01665e","#c7eae5",
                  "#8c510a","#f6e8c3")
plotdat %>% filter(!is.na(taxon.common)) %>%
ggplot() +
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`, col=site)) + 
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`), pch=21) +
  geom_rect(aes(xmin=9.5,xmax=8.5,ymin=1.75,ymax=2.25), col="gray60",fill=NA) +
  geom_segment(aes(x=9,xend=8.80,y=2.25,yend=2.4), col="gray60") +
  geom_text(data=data.frame(x=8.9,y=3.12,label="1 green crab"), aes(x=x,y=y,label=label), col="gray60", fontface="italic",size=4) +
  scale_color_manual(values=col_palette2,name="") +
  scale_size_continuous(range=c(2,6)) +
  guides(color=FALSE) +
  labs(x="",y="") + ggtitle("What bivalves, fish, and other crab are being eaten?") +
  coord_flip() + theme_bw() + theme(axis.text.x=element_text(angle=20,hjust=1,size=12),
                                    axis.text.y=element_text(size=12),
                                    legend.text=element_text(size=12), legend.title=element_text(size=13))
```
```{r}
png(here('figs','POSTER_prey-clams-crab-fish_by-site_dotplot.png'),res=300,height=1500,width=2000,bg = 'transparent')
plotdat %>% filter(!is.na(taxon.common)) %>%
ggplot() +
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`, col=site)) + 
  geom_point(aes(x=taxon.common,y=site,size=`Percent of Crabs\nwith food item`), pch=21) +
  geom_rect(aes(xmin=9.5,xmax=8.5,ymin=1.75,ymax=2.25), col="gray60",fill=NA) +
  geom_segment(aes(x=9,xend=8.80,y=2.25,yend=2.4), col="gray60") +
  geom_text(data=data.frame(x=8.9,y=3.32,label="1 green crab"), aes(x=x,y=y,label=label), col="gray60", fontface="italic",size=4) +
  scale_color_manual(values=col_palette2,name="") +
  scale_size_continuous(range=c(2,6)) +
  guides(color=FALSE) +
  labs(x="",y="") + ggtitle("What bivalves, fishes, and other\n crabs are being eaten?") +
  coord_flip() + theme_bw() + theme(axis.text.x=element_text(angle=30,hjust=1,size=12),
                                    axis.text.y=element_text(size=12),
                                    legend.text=element_text(size=12), legend.title=element_text(size=13),
                                    panel.grid.major=element_line(linewidth=0.25),
                                    # panel.background = element_rect(fill='transparent'), #transparent panel bg
                                    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                                    legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                                    legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
dev.off()
```



# Map

### Site Map
```{r}
willapa2 <- readRDS(here("data","metadata","willapa_terrain_background2.rds"))
willapa <- readRDS(here("data","metadata","willapa_terrain_background.rds"))
mapdat <- site_coords %>% dplyr::select(site_name,latitude,longitude,type) %>% distinct() %>%
  mutate(type=ifelse(type=="Sentinel","Monitoring\nSites","Clam Bed\nSites"))
```
```{r}
mymap <- ggmap(willapa2) + 
  geom_label_repel(data=mapdat, aes(x=longitude,y=latitude, label=site_name, color=type),
                   segment.colour = 'black',
                   nudge_x=0.2,
                   force=1, force_pull=0.2, segment.size=0.05, size=4) +
  geom_point(data=mapdat, aes(x=longitude,y=latitude, col=type), size=6) +
  geom_text(aes(x=-123.98,y=46.65,label="italic('Willapa\n Bay')"), col="deepskyblue4", size=4, parse=TRUE) +
  geom_text(aes(x=-123.87,y=46.23,label="italic('Columbia River')"), col="deepskyblue4", size=4, parse=TRUE) +
  ylab("Latitude") + xlab("Longitude") +
  scale_color_manual(values=c("#018571","#a6611a")) +
  theme(axis.title=element_blank(),
        axis.text=element_text(size=10),
        legend.title=element_blank(),
        legend.text=element_text(size=12), legend.position="right",legend.justification='top',
        plot.margin=unit(c(0,0,0,0), "cm")) +
  coord_fixed(ylim=c(46.3,46.8), xlim=c(-124.2, -123.7))
mymap
mymap2 <- ggmap(willapa) + 
  geom_label_repel(data=mapdat, aes(x=longitude,y=latitude, label=site_name, color=type),
                   segment.colour = 'black',
                   nudge_x=0.2,
                   force=1, force_pull=0.2, segment.size=0.05, size=5) +
  geom_point(data=mapdat, aes(x=longitude,y=latitude, col=type), size=6) +
  geom_text(aes(x=-123.98,y=46.65,label="italic('Willapa\n Bay')"), col="deepskyblue4", size=4, parse=TRUE) +
  geom_text(aes(x=-123.85,y=46.23,label="italic('Columbia River')"), col="deepskyblue4", size=4, parse=TRUE) +
  scale_color_manual(values=c("#018571","#a6611a")) +
  theme(axis.title=element_blank(),
        axis.text=element_blank(), axis.ticks=element_blank(),
        legend.title=element_blank(),
        legend.text=element_text(size=12), legend.position="right",legend.justification='top',
        plot.margin=unit(c(0.5,0,0,0), "cm"))
mymap2
```
```{r}
png(here('results','figures','fig1_map.png'),res=300, height=2000,width=1400)
mymap
dev.off()
png(here('results','figures','fig1_map_nolabels.png'),res=300, height=2000,width=1000)
mymap2
dev.off()
```


Alternative map
```{r}
require(raster)
pt.gadm <- getData('GADM', country='United States', level=0)
require(sf)
wc.lim = data.frame(ylim=c(44,48.3), xlim=c(-129, -119))
wc.bbox <- st_bbox(c(xmin=wc.lim$xlim[1],
                     xmax=wc.lim$xlim[2],
                     ymin=wc.lim$ylim[1],
                     ymax=wc.lim$ylim[2]))

pt.gadm <- sf::st_as_sf(pt.gadm) %>% 
  st_crop(wc.bbox)

loc.lim = data.frame(ylim=c(46.3,46.8), xlim=c(-124.2, -123.6))
loc.bbox <- st_bbox(c(
  xmin=loc.lim$xlim[1],
  xmax=loc.lim$xlim[2],
  ymin=loc.lim$ylim[1],
  ymax=loc.lim$ylim[2]))

# Altitude data (gives a raster list):
alt.gadm <- getData("alt", country = "United States")[[1]]
plot(alt.gadm, main="Elevation") 
slope.raster <- terrain(alt.gadm, opt='slope')
aspect.raster <- terrain(alt.gadm, opt='aspect')
hill.raster <- hillShade(slope.raster, aspect.raster, 40, 270)
hill.m <- rasterToPoints(hill.raster)
hill.df <-  data.frame(hill.m)
colnames(hill.df) <- c("lon", "lat", "hill")

```
```{r}
ggplot() +
  geom_sf(data=pt.gadm)+
  coord_sf(ylim=loc.lim$ylim, xlim=loc.lim$xlim)+
  geom_point(data=mapdat, aes(x=longitude,y=latitude, col=type), size=6) + 
  geom_raster(data = hill.df, aes(lon, lat, fill = hill), alpha = .45) +
  scale_fill_gradientn(colours = grey.colors(100)) +
  ggtitle("GADM border - local area") +
  theme_minimal()+
  theme(panel.background = element_rect(fill = 'lightblue'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```



### Carapace widths (mm)

```{r}
cw <- read_csv(here('data','metadata','Willapa Bay EGC Samples - Sample Data.csv')) %>%
  mutate(sequenced=ifelse(Sample_label %in% meta.run$sample, "Sequenced", "Trapped"))

cw %<>% mutate(site=ifelse(grepl("Stackpole",Site_Name), "Stackpole",
                                 ifelse(grepl("Oysterville",Site_Name), "Oysterville",
                                        ifelse(grepl("Nahcotta",Site_Name),"Nahcotta",
                                               ifelse(grepl("Long Beach",Site_Name),"Long Beach", NA))))) %>%
  mutate(site_type=ifelse(site %in% c("Stackpole","Nahcotta"),"Monitoring Sites","Clam Bed Sites"))

cw %>% group_by(sequenced, Sex) %>% summarise(ncrab=length(unique(Sample_label)))
```


```{r fig.width=6, fig.height=3}
cw.sequenced <- cw%>%filter(sequenced=="Sequenced")
plotdat <- cw %>% mutate(sequenced="Trapped") %>%
  bind_rows(cw.sequenced)

plotdat$site_type  <- factor(plotdat$site_type, levels=c("Clam Bed Sites","Monitoring Sites"))

ggplot(plotdat, aes(x=CW_mm, fill=sequenced)) +
  geom_histogram(position="dodge") +
  scale_fill_manual(values=c("black","grey60"),name="") +
  facet_grid(rows=vars(site_type),switch="y") + 
  scale_y_continuous(position = "right") +
  labs(x="Carapace Width (mm)",y="Number of Green Crabs") + theme_bw() + 
  theme(legend.position="top",strip.placement = "outside",
        strip.background=element_rect(color=c("white"), linewidth=2))
```

Alternative legend location
```{r}

ggplot(plotdat, aes(x=CW_mm, fill=sequenced)) +
  geom_histogram(position="dodge") +
  scale_fill_manual(values=c("black","grey60"),name="") +
  facet_grid(rows=vars(site_type),switch="y") + 
  scale_y_continuous(position = "right") +
  labs(x="Carapace Width (mm)",y="Number of Green Crabs") + theme_bw() + 
  theme(strip.placement = "outside",legend.position=c(0.1,0.9), legend.title=element_blank(),
        strip.background=element_rect(color=c("white"), linewidth=2))

plot1 <- ggplot(plotdat, aes(x=CW_mm, fill=sequenced)) +
  geom_histogram(position="dodge") +
  scale_fill_manual(values=c("black","grey60"),name="") +
  facet_grid(rows=vars(site_type),switch="y") + 
  scale_y_continuous(position = "right") +
  labs(x="Carapace Width (mm)",y="Number of Green Crabs") + theme_bw() + 
  theme(strip.placement = "outside",legend.position=c(0.4,0.9), legend.title=element_blank(),
        strip.text=element_text(size=12),
        strip.background=element_rect(color=c("white"), linewidth=2))
```


```{r}
png(here('results','figures','fig1_pamphlet_cw.png'),res=300, height=1000,width=1400)
plot1
dev.off()
```


### combined

```{r echo=TRUE}
right_side <- plot_grid(NULL, plot1, NULL, nrow=3, rel_heights=c(0.2,1,0.2))
png(here('results','figures','fig1-pamphet.png'),res=300, height=2300,width=2300)
plot_grid(mymap2,right_side,ncol=2, rel_widths=c(1,0.5))
dev.off()


png(here('results','figures','fig1_nolabels.png'),res=300, height=2000,width=2200)
plot_grid(mymap2,plot1,ncol=2, rel_widths=c(1,0.7))
dev.off()
```
